### 2021-10-28 theoriezf/Article.tex
Add files via upload

\documentclass[a4paper]{article}

\usepackage[top=2.5cm, bottom=2.5cm, left=2.5cm, right=2.5cm]{geometry}
\usepackage{ae,lmodern}
\usepackage[francais]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\begin{document}

\section{Introduction}

papier + dedukti + bout de théorie U (arrow, omicron)

\section{The language of pointed graphs}

\section{Formulas}

\section{Lemmas}

Expliquer qu'on n'a pas besoin des lemmes 1 et 2
Extensionnalité faible ajoutée entre 41 et 42 parce qu'elle est utilisée dans 44, 47, 48

\end{document}

### 2021-10-28 theoriezf/Article.tex
misc

@@
 \usepackage[T1]{fontenc}
+
+%\usepackage{paralist, tabularx}
+\usepackage{oldlfont,amssymb,epsf,stmaryrd,epsfig,color}\usepackage{amsmath}
+%\usepackage{hyperref}
+%\usepackage{framed,multicol,changepage}
+%\usepackage{graphicx}
+\usepackage[dvipsnames]{xcolor}
+%\usepackage{skull,ifoddpage,marginnote}
+%\usepackage{pifont}
+%\usepackage{graphicx}
+%\usepackage{ebproof}
+%\usepackage{mathtools}
+%\usepackage{bm}
+%\usepackage{stmaryrd}
+%\usepackage{cleveref}
+%\usepackage{tikz}
+%\usepackage{import} % For inkscape generated files
+
+\input{macros.tex}
 
@@
 
+\subsection{Sorts}
+
+The langage of the theory IZmod uses four sorts. The two first ones
+are for the pointed graphs and for the nodes of the pointed graphs.
+
+In Dedukti, we thus introduce two constants $\Graph$ and $\Node$ of
+type $\Type$
+\begin{verbatim}
+constant symbol Graph : TYPE;
+constant symbol Node : TYPE;
+\end{verbatim}
+As we have two sorts, we would need two universal quantifiers and two 
+existential quantifiers, one on each sort. 
+We rather use another solution \cite{theoryU} that is to declare a constant
+$\Set$ of type $\Type$ for codes of sorts,
+a function $\El$ of type $\Set \ra \Type$, two 
+$g$ and $n$ of sort $\Set$ and rewrite rules 
+$$\El~g \lra \Graph$$
+$$\El~n \lra \Node$$
+
+The two other sorts of the theory IZmod are for classes of nodes and
+for binary relations on nodes.
+
+In Dedukti, the sort of classes is just $\Node \ra \Prop$ and that of
+binary relations $\Node \ra \Node \ra \Prop$. To be able to quantify on 
+such sorts, we introduce constants $\omicron$ of type $\Set$ 
+$\arr $ of type
+$\Set \ra \Set \ra \Set$ 
+and rewrite rules
+$$El~\omicron \lra \Prop$$
+$$(El~(x \arr y)) \lra (El x) \ra (El y)$$
+
+Note that $El (n \arr \omicron) \lra \Node \ra \Prop$ and 
+$El (n \arr (n \arr \omicron)) \lra \Node \ra \Node \ra \Prop$.
+
 \section{Formulas}

### 2021-10-28 theoriezf/Article.tex
misc

--- 2021-10-28 theoriezf/Article.tex
@@
 
+%\usepackage{paralist, tabularx}
+\usepackage{oldlfont,amssymb,epsf,stmaryrd,epsfig,color}\usepackage{amsmath}
+%\usepackage{hyperref}
+%\usepackage{framed,multicol,changepage}
+%\usepackage{graphicx}
+\usepackage[dvipsnames]{xcolor}
+%\usepackage{skull,ifoddpage,marginnote}
+%\usepackage{pifont}
+%\usepackage{graphicx}
+%\usepackage{ebproof}
+%\usepackage{mathtools}
+%\usepackage{bm}
+%\usepackage{stmaryrd}
+%\usepackage{cleveref}
+%\usepackage{tikz}
+%\usepackage{import} % For inkscape generated files
+
+\input{macros.tex}
+
 \begin{document}
-
+\thispagestyle{empty}
 \section{Introduction}
@@
 
+\subsection{Sorts}
+
+The langage of the theory IZmod uses four sorts. The first two are for
+the pointed graphs and for the nodes of the pointed graphs.  In
+\dedukti, we thus introduce two constants $\Graph$ and $\Node$ of type
+$\Type$
+\begin{verbatim}
+constant symbol Graph : TYPE;
+constant symbol Node : TYPE;
+\end{verbatim}
+As we have two sorts, we would need two universal quantifiers and two
+existential quantifiers, one for each sort.  We rather use another
+solution \cite{theoryU} that is to declare a constant $\Set$ of type
+$\Type$ for codes of sorts, a function $\El$ of type $\Set \ra \Type$,
+two constants $g$ and $n$ of sort $\Set$ and rewrite rules
+$$\El~g \lra \Graph$$
+$$\El~n \lra \Node$$
+
+The two other sorts of the theory IZmod are for classes of nodes and
+for binary relations on nodes.  In \dedukti, the sort of classes is
+just $\Node \ra \Prop$ and that of binary relations
+$\Node \ra \Node \ra \Prop$. To quantify on such sorts, we introduce
+constants $\omicron$ of type $\Set$ and $\arr $ of type
+$\Set \ra \Set \ra \Set$ and rewrite rules
+$$\El~\omicron \lra \Prop$$
+$$\El~(x~\arr~y) \lra (\El~x) \ra (\El~y)$$
+Note that $\El~(n~\arr~\omicron)$ reduces to $\Node \ra \Prop$ and 
+$\El~(n~\arr~(n~\arr~\omicron))$ reduces to $\Node \ra \Node \ra \Prop$.
+
+The symbols $\Graph$, $\Node$, $g$ and $n$ are specific to the
+expression of $IZmod$ in \dedukti. In contrast the symbols $\Set$,
+$\El$, $\omicron$, and $\arr$ are part of the standard library of 
+\dedukti.
+
+\subsection{Signature}
+
+The signature of $IZmod$ contains 31 symbols. As we have replaced the
+sorts for classes and relations with the \dedukti types
+$\Node \ra \Prop$ and $\Node \ra \Node \ra \Prop$, we do not need
+specific predicate symbols to apply a class to a node or a relations
+to two. In the same way, we do not need comprehension axioms for 
+classes and relations.
+
+In the same way, the equality symbol is part of the standard library of 
+\dedukti.
+
+So the signature is reduced to 26 symbols.
+
+$$eta : \Graph \ra \Node \ra \Node \ra \Prop$$
+$$root : \Graph \ra \Node$$
+$$cr : \Graph \ra \Node \ra \Graph$$
+$$o : \Node$$
+$$O : \Node$$
+$$i : \Node \ra \Node$$
+$$i' : \Node \ra \Node$$
+$$j : \Node \ra \Node$$
+$$j' : \Node \ra \Node$$
+$$I : \Node \ra \Prop$$
+$$J : \Node \ra \Prop$$
+$$\rho : \Graph \ra \Node$$
+$$\hat{\rho} : \Node \ra \Graph$$
+$$Succ : \Node \ra \Node$$
+$$Pred : \Node \ra \Node$$
+$$Null : \Node \ra \Prop$$
+$$Nat : \Node \ra \Prop$$
+$$< : \Node \ra \Node \ra \Prop$$
+$$simeq : \Graph \ra \Graph \ra \Prop$$
+$$\in : \Graph \ra \Graph \ra \Prop$$
+$$join : \Graph \ra \Graph$$
+$$pair : \Graph \ra \Graph \ra \Graph$$
+$$powerset : \Graph \ra \Graph$$
+$$comp : \Graph \ra \El valuation \ra \El formula \ra \Graph$$
+$$omega : \Graph$$
+$$Cl : \Graph \ra \Graph$$
+
+À faire en verbatim plutôt qu'en math mode, mais un pb avec certains
+caractères unicode
+
 \section{Formulas}
+
+$$comp : \Graph \ra \El valuation \ra \El formula \ra \Graph$$
+
+In $comp$ we may have noticed the presence of valuation and formula. 
+Indeed.
+
+\section{The theory of pointed graphs}
 

### 2021-10-29 theoriezf/Article.tex
Add files via upload

@@
 \usepackage[T1]{fontenc}
+\usepackage{listings}
+\usepackage{amssymb}
+\usepackage{mathtools}
 
@@
 \input{macros.tex}
+\input{lambdapi.tex}
 
@@
 $\Type$
-\begin{verbatim}
+\begin{lstlisting}
 constant symbol Graph : TYPE;
 constant symbol Node : TYPE;
-\end{verbatim}
+\end{lstlisting}
 As we have two sorts, we would need two universal quantifiers and two
@@
 
-$$eta : \Graph \ra \Node \ra \Node \ra \Prop$$
-$$root : \Graph \ra \Node$$
-$$cr : \Graph \ra \Node \ra \Graph$$
-$$o : \Node$$
-$$O : \Node$$
-$$i : \Node \ra \Node$$
-$$i' : \Node \ra \Node$$
-$$j : \Node \ra \Node$$
-$$j' : \Node \ra \Node$$
-$$I : \Node \ra \Prop$$
-$$J : \Node \ra \Prop$$
-$$\rho : \Graph \ra \Node$$
-$$\hat{\rho} : \Node \ra \Graph$$
-$$Succ : \Node \ra \Node$$
-$$Pred : \Node \ra \Node$$
-$$Null : \Node \ra \Prop$$
-$$Nat : \Node \ra \Prop$$
-$$< : \Node \ra \Node \ra \Prop$$
-$$simeq : \Graph \ra \Graph \ra \Prop$$
-$$\in : \Graph \ra \Graph \ra \Prop$$
-$$join : \Graph \ra \Graph$$
-$$pair : \Graph \ra \Graph \ra \Graph$$
-$$powerset : \Graph \ra \Graph$$
-$$comp : \Graph \ra \El valuation \ra \El formula \ra \Graph$$
-$$omega : \Graph$$
-$$Cl : \Graph \ra \Graph$$
-
-À faire en verbatim plutôt qu'en math mode, mais un pb avec certains
-caractères unicode
+\begin{lstlisting}
+symbol eta : Graph → Node → Node → Prop;
+symbol root : Graph → Node;
+symbol cr : Graph → Node → Graph;
+constant symbol o : Node;
+constant symbol O : Node;
+symbol i : Node → Node;
+symbol i' : Node → Node;
+symbol j : Node → Node;
+symbol j' : Node → Node;
+symbol I : Node → Prop;
+symbol J : Node → Prop;
+symbol ρ : Graph → Node;
+symbol ρ^ : Node → Graph;
+symbol Succ : Node → Node;
+symbol Pred : Node → Node;
+symbol Null : Node → Prop;
+symbol Nat : Node → Prop;
+symbol < : Node → Node → Prop;
+symbol simeq : Graph → Graph → Prop;
+symbol ∈ : Graph → Graph → Prop;
+symbol join : Graph → Graph;
+symbol pair : Graph → Graph → Graph;
+symbol powerset : Graph → Graph;
+symbol comp : Graph → El valuation → El formula → Graph;
+symbol omega : Graph;
+symbol Cl : Graph → Graph;
+\end{lstlisting}
 

### 2021-10-29 theoriezf/lambdapi.tex
Add files via upload

\definecolor{green}{RGB}{0,130,0}
\definecolor{lightgrey}{RGB}{240,240,240}

\lstdefinelanguage{Lambdapi}
{
  inputencoding=utf8,
  extendedchars=true,
  numbers=none,
  numberstyle={},
  tabsize=2,
  basicstyle={\ttfamily\small\upshape},
  backgroundcolor=\color{white},
  keywords={abort,admit,admitted,apply,as,assert,assertnot,associative,assume,begin,builtin,commutative,compute,constant,debug,end,fail,flag,focus,generalize,have,in,induction,inductive,infix,injective,left,let,notation,off,on,opaque,open,prefix,print,private,proofterm,protected,prover,prover_timeout,quantifier,refine,reflexivity,require,rewrite,right,rule,sequential,simplify,solve,symbol,symmetry,type,TYPE,unif_rule,verbose,why3,with},
  sensitive=true,
  keywordstyle=\color{blue},
  morecomment=[l]{//},
  commentstyle={\itshape\color{red}},
  string=[b]{"},
  stringstyle=\color{orange},
  showstringspaces=false,
  literate=
  {↪}{$\hookrightarrow$}1
  {→}{$\rightarrow$}1
  {Π}{$\Pi$}1
  {≔}{$\coloneqq$}1
  {𝔹}{$\mathbb{B}$}1
  {𝕃}{$\mathbb{L}$}1
  {ℕ}{$\mathbb{N}$}1
  {α}{$\alpha$}1
  {π}{$\pi$}1
  {τ}{$\tau$}1
  {ω}{$\omega$}1
  {∧}{$\wedge$}1
  {≤}{$\le$}1
  {≠}{$\neq$}1
  {∉}{$\notin$}1
  {×}{$\times$}1
  {ρ}{$\rho$}1
  {∈}{$\in$}1 
}
\lstset{language={Lambdapi}}

### 2021-10-29 theoriezf/lambdapi.tex
Add files via upload

@@
+\usepackage{listings}
+\usepackage{amssymb}
+\usepackage{mathtools}
+\usepackage{xcolor}
 \definecolor{green}{RGB}{0,130,0}
@@
   tabsize=2,
-  basicstyle={\ttfamily\small\upshape},
-  backgroundcolor=\color{white},
+  basicstyle={\ttfamily\upshape\mdseries},
+  backgroundcolor=\color{lightgrey},
   keywords={abort,admit,admitted,apply,as,assert,assertnot,associative,assume,begin,builtin,commutative,compute,constant,debug,end,fail,flag,focus,generalize,have,in,induction,inductive,infix,injective,left,let,notation,off,on,opaque,open,prefix,print,private,proofterm,protected,prover,prover_timeout,quantifier,refine,reflexivity,require,rewrite,right,rule,sequential,simplify,solve,symbol,symmetry,type,TYPE,unif_rule,verbose,why3,with},

### 2021-11-14 theoriezf/Article.tex
Add files via upload
Update Article.tex

@@
 
-\usepackage[top=2.5cm, bottom=2.5cm, left=2.5cm, right=2.5cm]{geometry}
+\usepackage[top=3.5cm, bottom=3.5cm, left=3.5cm, right=3.5cm]{geometry}
 \usepackage{ae,lmodern}
-\usepackage[francais]{babel}
+\usepackage[english]{babel}
 \usepackage[utf8]{inputenc}
 \usepackage[T1]{fontenc}
-\usepackage{listings}
-\usepackage{amssymb}
-\usepackage{mathtools}
 
@@
 
+\title{Implementation of Set theory in \dedukti}
+\date{}
+
 \input{macros.tex}
@@
 \thispagestyle{empty}
+\maketitle
+
+\section*{Plan}
+
+Introduction : papier précédent, explications de la structure des graphes pointés, apparition de Dedukti depuis (passage de Deduction modulo à Lambpdapi calcul modulo), théorie U (arrow, omicron) \\
+
+The theory of pointed graphs : règles de réécritures \\
+
+The language of pointed graphs : types, signature \\
+
+Formulas : nécessité de restreindre l'ensemble des propositions dans la compréhension, définition des formules et des connecteurs sur les formules, principe de l'induction, principe de l'interprétation/valuation/update, présentations des nouveaux lemmes ajoutés pour prouver les résultats sur les formules (formulas-lemmas.lp) \\
+
+Lemmas : expliquer qu'on n'a pas besoin des lemmes 1 et 2, explication des mécanismes classiques réutilisés dans chaque preuve, détails de preuves pour certains lemmes intéressants, extensionalité faible ajoutée entre 41 et 42 parce qu'elle est utilisée dans 44, 47, 48 \\
+
+Conclusion : tableau récapitulatif du nombre de lignes de chaque preuve
+
 \section{Introduction}
@@
 
-\section{Formulas}
+\section{The language of formulas}
+\subsection{Formulas}
 
-$$comp : \Graph \ra \El valuation \ra \El formula \ra \Graph$$
+$$comp : \Graph \ra \El \ valuation \ra \El \ formula \ra \Graph$$
 
 In $comp$ we may have noticed the presence of valuation and formula. 
-Indeed.
+Indeed, the lemmas in which the $comp$ symbol is used are only valid for a subset of formula.
+The formulas need to have all of its quantifiers of sort $Graph$ and can only use the language $\in$, $\simeq$ and the classic logical connectives.
+
+Thus we need to introduce a set to caraterize the validity domain of such lemmas.
+In order to achieve this goal, we define the constant $formula$ of type $\Set$ and the logical connectives related to this constant.
+
+\begin{lstlisting}
+constant symbol formula : Set;
+constant symbol eqF : El nat → El nat → El formula;
+constant symbol inF : El nat → El nat → El formula;
+constant symbol andF : El formula → El formula → El formula;
+constant symbol orF : El formula → El formula → El formula;
+constant symbol allF : El nat → El formula → El formula;
+constant symbol exF : El nat → El formula → El formula;
+constant symbol impF : El formula → El formula → El formula;
+constant symbol fF : El formula;
+constant symbol tF : El formula;
+\end{lstlisting}
+
+\subsection{Interpretation}
+
+The next step is to interpret an object of type $formula$ into $Prop$.
+We introduce the constant $interpretation$ which receives a valuation of type $\El \ valuation$ and a formula of type $\El \ formula$ and return a $\Prop$.
+
+\begin{lstlisting}
+symbol interpretation : (El nat → Graph) → El formula → Prop;
+\end{lstlisting}
+
+The interpreation of formulas obeys the following rewriting rules :
+
+$$interpretation \ \sigma \ (eqF \ x \ y) \lra (\sigma \ x) \ \simeq \ (\sigma \ y)$$
+$$interpretation \ \sigma \ (inF \ x \ y) \lra (\sigma \ x) \ \in \ (\sigma \ y)$$
+$$interpretation \ \sigma \ (andF \ f \ g) \lra (interpretation \ \sigma \ f) \conj (interpretation \ \sigma \ g)$$
+$$interpretation \ \sigma \ (orF \ f \ g) \lra (interpretation \ \sigma \ f) \disj  (interpretation \sigma \ g)$$
+$$interpretation \ \sigma \ (impF \ f \ g) \lra (interpretation \ \sigma \ f) \imp (interpretation \ \sigma \ g)$$
+$$interpretation \ \sigma \ fF \lra \blbot$$
+$$interpretation \ \sigma \ tF \lra \bltop$$
+
+We need to have a tool to update a valuation when we assign a variable.
+To do so, we introduce the constant $update$ of type $(\El \ nat \ra Graph) \ra \El \ nat \ra Graph \ra (\El \ nat \ra Graph)$ which takes as arguments a valuation, a natural number and a graph and returns a new valuation.
+
++ A finir avec update1
++ supprimer El valuation ou pas ?
+
+\subsection{Theorems about valuation}
+
+\begin{theorem}
+$\fa \ \sigma, \fa \ x, y, z, \fa \ a, [(eqNP \ x \ y) \ \imp \ (((update1 \ \sigma \ x \ a) \ y \ z) \ \simeq \ a)] \ \conj \ [\blneg(eqNP \ x \ y) \ \imp \ (((update1 \ \sigma \ x \ a) \ y \ z) \ \simeq \ (\sigma \ z))]$
+\end{theorem}
+
+\begin{theorem}
+$\fa \ \sigma, \fa \ x, \fa \ a, b, (a \ \simeq \ b) \ \imp \ [\fa y, (update \ \sigma \ x \ a \ y) \ \simeq \ (update \ \sigma \ x \ b \ y)]$
+\end{theorem}
+
+\begin{theorem}
+$\fa \ \sigma, \fa \ x, y, \fa \ a, b, c, (a \ \simeq \ b) \ \imp \ (\fa z, (update \ (update \ \sigma \ x \ a) \ y \ c \ z) \ \simeq \ (update \  (update \ \sigma \ x \ b) \ y \ c \ z))$
+\end{theorem}
+
+\begin{theorem}
+$\fa \ \sigma, \sigma', \fa \ x, \fa \ c, (\fa y, \sigma \ y \ \simeq \ \sigma' \ y) \ \imp \ (\fa z, (update \ \sigma \ x \ c \ z) \ \simeq \ (update \ \sigma' \ x \ c \ z))$
+\end{theorem}
+
+\begin{theorem}
+$\fa \ f, \fa \ \sigma, \sigma', (interpretation \ \sigma \ f) \ \conj \ (\fa x, (\sigma \ x) \ \simeq \ (\sigma' \ x))) \ \imp \ (interpretation \ \sigma' \ f)$
+\end{theorem}
 
@@
 
-Expliquer qu'on n'a pas besoin des lemmes 1 et 2
-Extensionnalité faible ajoutée entre 41 et 42 parce qu'elle est utilisée dans 44, 47, 48
+\section{Conclusion}
+
+\section*{References}
 

### 2021-11-15 theoriezf/Article.tex
Add files via upload

@@
 In $comp$ we may have noticed the presence of valuation and formula. 
-Indeed, the lemmas in which the $comp$ symbol is used are only valid for a subset of formula.
+Indeed, the lemmas in which the $comp$ symbol is used are only valid for a subset of formulas.
 The formulas need to have all of its quantifiers of sort $Graph$ and can only use the language $\in$, $\simeq$ and the classic logical connectives.
@@
 constant symbol tF : El formula;
+\end{lstlisting}
+
+Then we are able to define an induction over formulas, using the language $\in$, $\simeq$ and the classic logical connectives.
+
+\begin{lstlisting}
+constant symbol recF : Π (P : El formula → Prop), 
+π(`∀ x, `∀ y, P (eqF x y))
+→ π(`∀ x, `∀  y, P (inF x y))
+→ π(`∀ f, `∀ g, (P f ∧ P g) ⇒ (P (andF f g)))
+→ π(`∀ f, `∀ g, (P f ∧ P g) ⇒ (P (orF f g)))
+→ π(`∀ f, `∀ g, (P f ∧ P g) ⇒ (P (impF f g)))
+→ π(`∀ f, (P f) ⇒ (`∀ x, P (allF x f)))
+→ π(`∀ f, (P f) ⇒ (`∀ x, P (exF x f)))
+→ π(P tF)
+→ π(P fF)
+→ π(`∀ f, P f);
 \end{lstlisting}
@@
 
+Thanks to the introduction of $interpretation$, we can deduce five theorems. \\
+
+The first theorem is used to simplify the terms when updating a valuation.
+
 \begin{theorem}
@@
 \end{theorem}
+
+The first term of the conjonction is proved by simple recurrence over natural numbers. The second term of the conjonction is proved by double recurrence. \\
+
+The second theorem conveys the idea that if two graphs are bisimilar then it is identical to update a valuation by either of theses two graphs.
 

### 2021-11-15 theoriezf/lambdapi.tex
Add files via upload

@@
   {∈}{$\in$}1
+  {∀}{$\forall$}1
+  {⇒}{$\Rightarrow$}1  
 }

### 2021-11-15 theoriezf/Article.tex
Update Article.tex

@@
 
+The language of pointed graphs : types, signature \\
+
 The theory of pointed graphs : règles de réécritures \\
-
-The language of pointed graphs : types, signature \\
 
@@
 
+We define the symbol $update1$ of type $(\El \ nat \ \ra \ Graph) \ \ra \ \El \ nat \ \ra \ Graph \ \ra \ \El \ nat \ \ra \ (\El \ nat \ \ra \ Graph)$ which obeys the following rewriting rules :
+
+$$update \ \sigma \ x \ a \ y \ \lra \ update1 \ \sigma \ x \ a \ y \ y$$
+$$update \ \sigma \ zero \ a \ zero \ z \ \lra \ a$$
+$$update \ \sigma \ zero \ a \ (s \ y) \ z \ \lra \ \sigma \ z$$
+$$update \ \sigma \ (s \ x) \ a \ zero \ z \ \lra \ \sigma \ z$$
+$$update \ \sigma \ (s \ x) \ a \ (s \ y) \ z \ \lra \ update1 \ \sigma \ x \ a \ y \ z$$
+
 + A finir avec update1
@@
 
-\subsection{Theorems about valuation}
+\subsection{Results concerning valuation}
 
@@
 
-The first term of the conjonction is proved by simple recurrence over natural numbers. The second term of the conjonction is proved by double recurrence. \\
+\begin{proof}
+The first term of the conjonction is proved by simple recurrence over natural numbers. The second term of the conjonction is proved by double recurrence.
+\end{proof}
 
@@
 
+The fourth theorem states that if two valuations are equal they keep being equal after an update.
+
 \begin{theorem}
@@
 
+\begin{proof}
+The fifth theorem is proved by induction over formulas.
+\end{proof}
+
 \section{The theory of pointed graphs}
@@
 
+\begin{center}
+\begin{tabular}{|c|c||c|c|}
+\hline Lemma & Number of lines in the proof & Lemma & Number of lines in the proof \\
+\hline 3 & ? & 29 & ? \\
+\hline 4 & ? & 30 & ? \\
+\hline 5 & ? & 31 & ? \\
+\hline 6 & ? & 32 & ? \\
+\hline 7 & ? & 33 & ? \\
+\hline 8 & ? & 34 & ? \\
+\hline 9 & ? & 35 & ? \\
+\hline 10 & ? & 36 & ? \\
+\hline 11 & ? & 37 & ? \\
+\hline 12 & ? & 38 & ? \\
+\hline 13 & ? & 39 & ? \\
+\hline 14 & ? & 40 & ? \\
+\hline 15 & ? & 41 & ? \\
+\hline 16 & ? & 42 & ? \\
+\hline 17 & ? & 43 & ? \\
+\hline 18 & ? & 44 & ? \\
+\hline 19 & ? & 45 & ? \\
+\hline 20 & ? & 46 & ? \\
+\hline 21 & ? & 47 & ? \\
+\hline 22 & ? & 48 & ? \\
+\hline 23 & ? & 49 & ? \\
+\hline 24 & ? & 50 & ? \\
+\hline 25 & ? & 51 & ? \\
+\hline 26 & ? & 52 & ? \\
+\hline 27 & ? & 53 & ? \\
+\hline 28 & ? & & \\
+\hline
+\end{tabular}
+\end{center}
+
 \section*{References}

### 2021-11-17 theoriezf/lambdapi.tex
Add files via upload

--- 2021-10-29 theoriezf/lambdapi.tex
@@
   {α}{$\alpha$}1
+  {λ}{$\lambda$}1
+  {σ}{$\sigma$}1
   {π}{$\pi$}1
@@
   {∧}{$\wedge$}1
+  {∨}{$\vee$}1
   {≤}{$\le$}1
@@
   {∈}{$\in$}1
+  {∀}{$\forall$}1
+  {∃}{$\exists$}1
+  {⇒}{$\Rightarrow$}1
+  {⇔}{$\Leftrightarrow$}1
 }

### 2021-11-17 theoriezf/Article/Article.tex
Add files via upload

--- 2021-11-15 theoriezf/Article.tex
@@
 %\usepackage{import} % For inkscape generated files
+\usepackage{biblatex}
+\usepackage{url}
 
@@
 \input{lambdapi.tex}
+\addbibresource{ref.bib}
 
@@
 
-Introduction : papier précédent, explications de la structure des graphes pointés, apparition de Dedukti depuis (passage de Deduction modulo à Lambpdapi calcul modulo), théorie U (arrow, omicron) \\
+Introduction : introduction du papier précédent, apparition de Dedukti depuis (passage de Deduction modulo à Lambpdapi calcul modulo), théorie U (arrow, omicron) \\
 
-The language of pointed graphs : types, signature \\
+The theory of pointed graphs :  explications détaillées de la structure des graphes pointés \\
 
-The theory of pointed graphs : règles de réécritures \\
+The language of pointed graphs : types, signature, règles de réécritures \\
 
@@
 
-papier + dedukti + bout de théorie U (arrow, omicron)
+\section{The theory of pointed graphs}
 
@@
 symbol powerset : Graph → Graph;
-symbol comp : Graph → El valuation → El formula → Graph;
+symbol comp : Graph → (El nat → Graph) → El formula → Graph;
 symbol omega : Graph;
@@
 
+\subsection{Rewriting rules}
+
 \section{The language of formulas}
-\subsection{Formulas}
 
-$$comp : \Graph \ra \El \ valuation \ra \El \ formula \ra \Graph$$
-
-In $comp$ we may have noticed the presence of valuation and formula. 
-Indeed, the lemmas in which the $comp$ symbol is used are only valid for a subset of formulas.
+We may have notice that the lemmas in which the $comp$ symbol is used are only valid for a subset of formulas.
 The formulas need to have all of its quantifiers of sort $Graph$ and can only use the language $\in$, $\simeq$ and the classic logical connectives.
@@
 Thus we need to introduce a set to caraterize the validity domain of such lemmas.
+
+\subsection{Formulas}
+
 In order to achieve this goal, we define the constant $formula$ of type $\Set$ and the logical connectives related to this constant.
@@
 \end{lstlisting}
+
+Henceforth, we are able to define in \dedukti
+$$comp : \Graph \ra (\El \ nat \ \ra \ Graph) \  \ra \El \ formula \ra \Graph$$
 
@@
 The next step is to interpret an object of type $formula$ into $Prop$.
-We introduce the constant $interpretation$ which receives a valuation of type $\El \ valuation$ and a formula of type $\El \ formula$ and return a $\Prop$.
+We introduce the constant $interpretation$ which receives a valuation of type $\El \ nat \ \ra \ Graph$ and a formula of type $\El \ formula$ and return a $\Prop$.
 
@@
 
-The interpreation of formulas obeys the following rewriting rules :
+We need to have a tool to update a valuation when we assign a variable.
+To do so, we introduce the constant $update$ of type $(\El \ nat \ra Graph) \ra \El \ nat \ra Graph \ra (\El \ nat \ra Graph)$ which takes as arguments a valuation, a natural number and a graph and returns a new valuation: $(update \ \sigma \ x \ a)$ is a new valuation that substitues $x$ by $a$ and acts like $\sigma$ for the other natural numbers.
+
+To write a rewriting rule upon $update$, we need to be able to check if we apply $(update \ \sigma \ x \ a)$ to $z = x$ or to $z \neq x$.
+
+We define the symbol $update1$ of type $(\El \ nat \ \ra \ Graph) \ \ra \ \El \ nat \ \ra \ Graph \ \ra \ \El \ nat \ \ra \ (\El \ nat \ \ra \ Graph)$.
+We keep in memory the argument of the valuation with the rewriting rule:
+
+$$update \ \sigma \ x \ a \ y \ \lra \ update1 \ \sigma \ x \ a \ y \ y$$
+
+We have two natural numbers to compare: $x$, which is substitued by $a$ by the $update$, and $y$ that is the argument we apply to the valuation.
+The technique we use to compare $x$ and $y$ is the following: we keep in memory $y$ in the variable $z$, we decrement the two numbers until either one or both are equal to zero.
+If both are equal to zero, then $x$ and $y$ are equal and we return $a$.
+If only one equals zero, then they are different and we return $\sigma \ z$.
+
+$$update1 \ \sigma \ (s \ x) \ a \ (s \ y) \ z \ \lra \ update1 \ \sigma \ x \ a \ y \ z$$
+$$ update1 \ \sigma \ zero \ a \ zero \ z \ \lra \ a$$
+$$update1 \ \sigma \ zero \ a \ (s \ y) \ z \ \lra \ \sigma \ z$$
+$$update1 \ \sigma \ (s \ x) \ a \ zero \ z \ \lra \ \sigma \ z$$
+
+Now we have all the tools to define the rewiting rules of the interpreation of formulas:
 
@@
 $$interpretation \ \sigma \ (impF \ f \ g) \lra (interpretation \ \sigma \ f) \imp (interpretation \ \sigma \ g)$$
+$$interpretation \ \sigma \ (allF \ x \ f) \ \lra \ \fa \ a, interpretation \ (update \ \sigma \ x \ a) \ f$$
+$$interpretation \ \sigma \ (exF \ x \ f) \ \lra \ \ex \ a, interpretation \ (update \ \sigma \ x \ a) \ f$$
 $$interpretation \ \sigma \ fF \lra \blbot$$
 $$interpretation \ \sigma \ tF \lra \bltop$$
-
-We need to have a tool to update a valuation when we assign a variable.
-To do so, we introduce the constant $update$ of type $(\El \ nat \ra Graph) \ra \El \ nat \ra Graph \ra (\El \ nat \ra Graph)$ which takes as arguments a valuation, a natural number and a graph and returns a new valuation.
-
-We define the symbol $update1$ of type $(\El \ nat \ \ra \ Graph) \ \ra \ \El \ nat \ \ra \ Graph \ \ra \ \El \ nat \ \ra \ (\El \ nat \ \ra \ Graph)$ which obeys the following rewriting rules :
-
-$$update \ \sigma \ x \ a \ y \ \lra \ update1 \ \sigma \ x \ a \ y \ y$$
-$$update \ \sigma \ zero \ a \ zero \ z \ \lra \ a$$
-$$update \ \sigma \ zero \ a \ (s \ y) \ z \ \lra \ \sigma \ z$$
-$$update \ \sigma \ (s \ x) \ a \ zero \ z \ \lra \ \sigma \ z$$
-$$update \ \sigma \ (s \ x) \ a \ (s \ y) \ z \ \lra \ update1 \ \sigma \ x \ a \ y \ z$$
-
-+ A finir avec update1
-+ supprimer El valuation ou pas ?
 
@@
 
-\section{The theory of pointed graphs}
+\section{Lemmas}
 
-\section{Lemmas}
+The first lemma $x=x$ does not need to be implemented since it is already part of the standard library of \dedukti \ under the name $refl$ (which is polymorphic).
+
+The second lemma is not required as it is a consequence of the rewriting rule of the polumorphic $=$ implemented in the stardard library of \dedukti: 
+\begin{lstlisting}
+constant symbol = {s} : El s → El s → Prop;
+notation = infix 4;
+rule π (@= $s $x $y) ↪ Π (P : El $s → Prop), π(P $x) → π(P $y);
+\end{lstlisting}
+
+\subsection{Useful mechanisms}
+
+All the other lemmas except the ones where $comp$ is involved are proved using the blueprint in \cite{zermodulo53}.
+The complete proofs can be found in \url{https://github.com/ttraversie/zf/tree/main/theoriezf}.
+
+\subsection{Some interesting proofs}
+
+Lemma32
+
+\subsection{Weak extensionnality}
+
+We want to deduce $weak \ extensionality$ from $strong \ extensionality$ (i.e. lemma 41).
+
+\begin{lstlisting}
+opaque symbol lemma41 : Π (x y : El nat), Π (f : El formula), Π (c d : Graph), π(`∀ σ : (El nat → Graph), ((interpretation (update (update σ x c) y d) f)
+∧ (`∀ a, `∀ a', `∀ b, ((a' ∈ a) ∧ (interpretation (update (update σ x a) y b) f)) ⇒ (`∃ b', ((b' ∈ b) ∧ (interpretation (update (update σ x a') y b') f))))
+∧ (`∀ b, `∀ b', `∀ a, ((b' ∈ b) ∧ (interpretation (update (update σ x a) y b) f)) ⇒ (`∃ a', ((a' ∈ a) ∧ (interpretation (update (update σ x a') y b') f))))) ⇒ (c simeq d))
+
+opaque symbol lemmaExt : Π (c d : Graph), π((`∀ x, (x ∈ c) ⇔ (x ∈ d)) ⇒ (c simeq d))
+\end{lstlisting}
+
+To do this, we need to prove the following intermediry lemma:
+
+\begin{lstlisting}
+opaque symbol lemmaHypExt : Π (c d : Graph), π((`∀ z, (z ∈ c) ⇔ (z ∈ d)) ⇒ ((((c simeq c) ∧ (d simeq d)) ∨ (c simeq d))
+∧ (`∀ a, `∀ a', `∀ b, ((a' ∈ a) ∧ (((a simeq c) ∧ (b simeq d)) ∨ (a simeq b))) ⇒ (`∃ b', ((b' ∈ b) ∧ (((a' simeq c) ∧ (b' simeq d)) ∨ (a' simeq b')))))
+∧ (`∀ b, `∀ b', `∀ a, ((b' ∈ b) ∧ (((a simeq c) ∧ (b simeq d)) ∨ (a simeq b))) ⇒ (`∃ a', ((a' ∈ a) ∧ (((a' simeq c) ∧ (b' simeq d)) ∨ (a' simeq b')))))))
+\end{lstlisting}
+
+We thus can prove $weak \ extensionality$ thanks to:
+
+\begin{lstlisting}
+opaque symbol lemmaExt : Π (c d : Graph), π((`∀ x, (x ∈ c) ⇔ (x ∈ d)) ⇒ (c simeq d))
+≔ begin
+assume c d H;
+refine lemma41 zero one (orF (andF (eqF zero two) (eqF one three)) (eqF zero one)) c d (update (update (λ _, empty_set) two c) three d) (lemmaHypExt c d H);
+end;
+\end{lstlisting}
 
@@
 \hline 27 & ? & 53 & ? \\
-\hline 28 & ? & & \\
+\hline 28 & ? & Weak extensionality & ? \\
 \hline
@@
 
-\section*{References}
+\printbibliography
 

### 2021-12-02 theoriezf/Article/Article.tex
Add files via upload

@@
 %\usepackage{cleveref}
-%\usepackage{tikz}
+\usepackage{tikz}
 %\usepackage{import} % For inkscape generated files
@@
 
-\title{Implementation of Set theory in \dedukti}
+\title{Title}
 \date{}
@@
 
-\section*{Plan}
-
-Introduction : introduction du papier précédent, apparition de Dedukti depuis (passage de Deduction modulo à Lambpdapi calcul modulo), théorie U (arrow, omicron) \\
-
-The theory of pointed graphs :  explications détaillées de la structure des graphes pointés \\
-
-The language of pointed graphs : types, signature, règles de réécritures \\
-
-Formulas : nécessité de restreindre l'ensemble des propositions dans la compréhension, définition des formules et des connecteurs sur les formules, principe de l'induction, principe de l'interprétation/valuation/update, présentations des nouveaux lemmes ajoutés pour prouver les résultats sur les formules (formulas-lemmas.lp) \\
-
-Lemmas : expliquer qu'on n'a pas besoin des lemmes 1 et 2, explication des mécanismes classiques réutilisés dans chaque preuve, détails de preuves pour certains lemmes intéressants, extensionalité faible ajoutée entre 41 et 42 parce qu'elle est utilisée dans 44, 47, 48 \\
-
-Conclusion : tableau récapitulatif du nombre de lignes de chaque preuve
-
 \section{Introduction}
 
+\dedukti \ is a logical framework based upon the language \lpcm, which was created to overcome the lack of interoperability between proof assistanrs.
+As it is a logical framework, we can implement in \dedukti \ theories.
+Especially, \textit{Set theory} is interesting beacause it will enable traductions from and towards \textsc{Isabelle/ZF}.
+
+A previous work \cite{zermodulo} has demonstrate the implemtation of \textit{Set theory} in \textit{Deduction modulo} thaks to a structure of \textit{pointed graphs}.
+Until the writing of \cite{zermodulo}, \lpcm \ has been created as an improvement of \textit{Deduction modulo}.
+We adapt the implementation to the \lpcm \ and then implement it in \dedukti.
+
 \section{The theory of pointed graphs}
+
+To implement $Set \ theory$ in \dedukti, the first idea is to translate the axioms of the theory into rewriting rules.
+Nevertheless, as pointed out by M.
+Crabbé \cite{crabbé}, such implementation does not guarantee normalization.
+That is why A.
+Miquel and G.
+Dowek \cite{zermodulo} introduced a theory of pointed graphs such that the implementation of $Set~theory$ respects normalization.
+
+\subsection{Pointed graphs}
+
+Sets are reprented by pointed graphs. A pointed graph is characterizd by a direct graph and one of its node identified as the root.
+
+\begin{figure}[h]
+\begin{tikzpicture}[>=stealth,yscale=1.8,xscale=1.4]
+
+\node(a) at (3,3)[circle,draw,text centered] {a};
+\node(b) at (5,3)[circle,draw,text centered] {b};
+\node(c) at (5,2.5)[circle,draw,text centered] {c};
+\node(d) at (7,3)[circle,draw,text centered] {d};
+\node(e) at (6.5,2.5)[circle,draw,text centered] {e};
+\node(f) at (7.5,2.5)[circle,draw,text centered] {f};
+\node(g) at (7.5,2)[circle,draw,text centered] {g};
+
+\draw[->] (b) -- (c); 
+\draw[->] (d) -- (e);
+\draw[->] (d) -- (f);
+\draw[->] (f) -- (g);
+
+\end{tikzpicture}
+\end{figure}
+
+The first graph with root $a$ represents the set $\emptyset$.
+The second graph with root $b$ represents the set $\{\emptyset\}$, while with root c it represents $\emptyset$.
+The third graph with root $d$ represents $\{\emptyset,\{\emptyset\}\}$, \textit{etc}.
+
+The operator $root$ gives the root of a pointed graph.
+$a/x$ replace the root of $a$ by $x$.
+$x \eta_a y$ means there is a edge in $a$ from $y$ to $x$.
+$rel(x,y,r)$ means that $x$ and $y$ are related by the relation $r$.
+
+Because some different pointed graphs represents the same set, we introduce the bisimilarity:
+
+$a \simeq b \lra \ex r~(rel(root(a),root(b),r) \\
+\conj \ \fa x \fa x'~\fa y~(x'~\eta_a~x~\conj~rel(x, y, r)~\imp~\ex~y'~(y'~\eta_b~y~\conj~rel(x', y', r))) \\
+\conj \ \fa y \fa y'~\fa x~(y'~\eta_b~y~\conj~rel(x, y, r)~\imp~\ex~x'~(x'~\eta_a~x~\conj~rel(x', y', r))))$
+
+\subsection{The theory IZmod}
+
+IZmod
 
@@
 the pointed graphs and for the nodes of the pointed graphs.  In
-\dedukti, we thus introduce two constants $\Graph$ and $\Node$ of type
-$\Type$
-\begin{lstlisting}
-constant symbol Graph : TYPE;
-constant symbol Node : TYPE;
-\end{lstlisting}
-As we have two sorts, we would need two universal quantifiers and two
+\dedukti, we would need two universal quantifiers and two
 existential quantifiers, one for each sort.  We rather use another
@@
 $\Type$ for codes of sorts, a function $\El$ of type $\Set \ra \Type$,
-two constants $g$ and $n$ of sort $\Set$ and rewrite rules
-$$\El~g \lra \Graph$$
-$$\El~n \lra \Node$$
+two constants $graph$ and $node$ of sort $\Set$.
+
+\begin{lstlisting}
+constant symbol graph : Set;
+constant symbol node : Set;
+\end{lstlisting}
 
@@
 for binary relations on nodes.  In \dedukti, the sort of classes is
-just $\Node \ra \Prop$ and that of binary relations
-$\Node \ra \Node \ra \Prop$. To quantify on such sorts, we introduce
-constants $\omicron$ of type $\Set$ and $\arr $ of type
-$\Set \ra \Set \ra \Set$ and rewrite rules
-$$\El~\omicron \lra \Prop$$
+just $\El~node \ra \El~prop$ and that of binary relations
+$\El~node \ra \El~node \ra \El~prop$. To quantify on such sorts, we introduce constant $\arr $ of type
+$\Set \ra \Set \ra \Set$ and rewrite rule
 $$\El~(x~\arr~y) \lra (\El~x) \ra (\El~y)$$
-Note that $\El~(n~\arr~\omicron)$ reduces to $\Node \ra \Prop$ and 
-$\El~(n~\arr~(n~\arr~\omicron))$ reduces to $\Node \ra \Node \ra \Prop$.
 
-The symbols $\Graph$, $\Node$, $g$ and $n$ are specific to the
+The symbols $graph$ and $node$ are specific to the
 expression of $IZmod$ in \dedukti. In contrast the symbols $\Set$,
-$\El$, $\omicron$, and $\arr$ are part of the standard library of 
+$\El$, $prop$, and $\arr$ are part of the standard library of 
 \dedukti.
@@
 The signature of $IZmod$ contains 31 symbols. As we have replaced the
-sorts for classes and relations with the \dedukti types
-$\Node \ra \Prop$ and $\Node \ra \Node \ra \Prop$, we do not need
-specific predicate symbols to apply a class to a node or a relations
-to two. In the same way, we do not need comprehension axioms for 
-classes and relations.
+sorts for classes and relations with the \dedukti \ types
+$\El~node \ra \El~prop$ and $\El~node \ra \El~node \ra \El~prop$, we do not need specific predicate symbols to apply a class to a node or a relations to two.
+In the same way, we do not need comprehension axioms for classes and relations.
 
-In the same way, the equality symbol is part of the standard library of 
-\dedukti.
+Similarly, the equality symbol is part of the standard library of \dedukti.
 
-So the signature is reduced to 26 symbols.
+So the signature is reduced to 26 symbols. The specific case of the comprehension symbol is treated later.
 
 \begin{lstlisting}
-symbol eta : Graph → Node → Node → Prop;
-symbol root : Graph → Node;
-symbol cr : Graph → Node → Graph;
-constant symbol o : Node;
-constant symbol O : Node;
-symbol i : Node → Node;
-symbol i' : Node → Node;
-symbol j : Node → Node;
-symbol j' : Node → Node;
-symbol I : Node → Prop;
-symbol J : Node → Prop;
-symbol ρ : Graph → Node;
-symbol ρ^ : Node → Graph;
-symbol Succ : Node → Node;
-symbol Pred : Node → Node;
-symbol Null : Node → Prop;
-symbol Nat : Node → Prop;
-symbol < : Node → Node → Prop;
-symbol simeq : Graph → Graph → Prop;
-symbol ∈ : Graph → Graph → Prop;
-symbol join : Graph → Graph;
-symbol pair : Graph → Graph → Graph;
-symbol powerset : Graph → Graph;
-symbol comp : Graph → (El nat → Graph) → El formula → Graph;
-symbol omega : Graph;
-symbol Cl : Graph → Graph;
+symbol eta : El graph → El node → El node → El prop;
+symbol root : El graph → El node;
+symbol cr : El graph → El node → El graph;
+constant symbol o : El node;
+constant symbol O : El node;
+symbol i : El node → El node;
+symbol i' : El node → El node;
+symbol j : El node → El node;
+symbol j' : El node → El node;
+symbol I : El node → El prop;
+symbol J : El node → El prop;
+symbol ρ : El graph → El node;
+symbol ρ^ : El node → El graph;
+symbol Succ : El node → El node;
+symbol Pred : El node → El node;
+symbol Null : El node → El prop;
+symbol Nat : El node → El prop;
+symbol < : El node → El node → El prop;
+symbol simeq : El graph → El graph → El prop;
+symbol ∈ : El graph → El graph → El prop;
+symbol join : El graph → El graph;
+symbol pair : El graph → El graph → El graph;
+symbol powerset : El graph → El graph;
+symbol omega : El graph;
+symbol Cl : El graph → El graph;
 \end{lstlisting}
 
+\subsection{From \textit{Deduction modulo} to \lpcm}
+
+Since the publication of \cite{zermodulo}, \lpcm has been developped as an evolution of \textit{Deduction modulo}.
+
+More specifically, \lpcm allows to quantify on proposition.
+
+In \dedukti, the symbol $mem$ and $rel$ are useless. Indeed, we can apply a node $x$ to class $P$ with $P~x$ and two nodes $x$ and $y$ to a relation $r$ with $r~x~y$.
+
 \subsection{Rewriting rules}
+
+The rewriting rule $$y = z \ \lra \ \fa \ p \ (mem(y, p) \ \imp \ mem(z, p))$$ is not required as it is a consequence of the rewriting rule of the polymorphic $=$ implemented in the stardard library of \dedukti:
+\begin{lstlisting}
+constant symbol = [s] : El s → El s → El prop;
+notation = infix 4;
+rule π (@= $s $x $y) ↪ Π (P : El $s → El prop), π(P $x) → π(P $y);
+\end{lstlisting}
+
++ Expliquer pourquoi on n'utilise pas les symboles g et g' \\
+
+All the other rewriting rules, except the ones involving $comp$, are easy to implement in \dedukti.
+
+\begin{lstlisting}
+rule eta (cr $a $z) $x $y ↪ eta $a $x $y;
+rule root (cr $a $x) ↪ $x;
+rule (cr (cr $a $x) $y) ↪ cr $a $y;
+
+rule i' (i $x) ↪ $x;
+rule j' (j $x) ↪ $x;
+rule ρ^ (ρ $x) ↪ $x;
+rule I (i $x) ↪ ⊤;
+rule J (j $x) ↪ ⊤;
+rule I (j $x) ↪ ⊥;
+rule J (i $x) ↪ ⊥;
+rule I (o) ↪ ⊥;
+rule J (o) ↪ ⊥;
+
+rule Pred (Succ $x) ↪ $x ;
+rule Null O ↪ ⊤;
+rule Nat O ↪ ⊤;
+rule Null (Succ $x) ↪ ⊥;
+rule Nat (Succ $x) ↪ Nat $x;
+
+rule $x < O ↪ ⊥;
+rule $x < (Succ $y) ↪ ($x < $y) ∨ ($x = $y);
+
+rule $a simeq $b ↪ `∃ r : El (node arrow (node arrow prop)), 
+    r (root $a) (root $b)
+    ∧ (`∀ x, `∀ x', `∀ y, 
+        eta $a x' x ∧ r x y
+            ⇒ `∃ y', eta $b y' y ∧ r x' y')
+    ∧ (`∀ y, `∀ y', `∀ x, 
+        eta $b y' y ∧ r x y
+            ⇒ `∃ x', eta $a x' x ∧ r x' y');
+
+rule $a ∈ $b ↪ `∃ x, ((eta $b x (root $b)) ∧ ($a simeq cr $b x));
+
+rule eta (join $a) $x $x' ↪ 
+	(`∃ y, `∃ y', ($x = i y) ∧ ($x' = i y') ∧ eta $a y y')
+    ∨ (`∃ y, `∃ z, ($x = i y) 
+    	∧ ($x' = o) 
+    	∧ eta $a y z 
+    	∧ eta $a z (root $a));
+
+rule eta (pair $a $b) $x $x' ↪ 
+	(`∃ y, `∃ y', (($x = i y) ∧ ($x' = i y') ∧ eta $a y y'))
+    ∨ (`∃ y, `∃ y', ($x = j y) ∧ ($x' = j y') ∧ eta $b y y')
+    ∨ (($x = i (root $a)) ∧ ($x' = o))
+    ∨ (($x = j (root $b)) ∧ ($x' = o));
+
+rule eta (powerset $a) $x $x' ↪ 
+	(`∃ y, `∃ y', ($x = i y) ∧ ($x' = i y') ∧ eta $a y y')
+    ∨ (`∃ y, `∃ c, ($x = i y) 
+    	∧ ($x' = j (ρ c)) 
+    	∧ (eta $a y (root $a)) 
+    	∧ ((cr $a y) ∈ c))
+    ∨ (`∃ c, ($x = j (ρ c)) ∧ ($x' = o));
+
+symbol omega : El graph;
+rule eta omega $x $x' ↪ 
+	(`∃ y, `∃ y', ($x = i y) ∧ ($x' = i y') ∧ (y < y'))
+    ∨ (`∃ y, ($x = i y) ∧ ($x' = o) ∧ Nat y);
+
+rule eta (Cl $a) $x $x' ↪ 
+	(`∃ y, `∃ y', (($x = i y) ∧ ($x' = i y') ∧ eta $a y y'))
+    ∨ (`∃ y, ($x = i y) 
+        ∧ ($x' = o)
+        ∧ (`∀ c : El (node arrow prop), 
+                ((`∀ z, eta $a z (root $a) ⇒ c z)
+                ∧ (`∀ z, `∀ z', (eta $a z z') ∧ (c z') ⇒ (c z)))
+            ⇒ c y));
+
+rule root (join $a) ↪ o;
+rule root (pair $a $b) ↪ o;
+rule root (powerset $a) ↪ o;
+rule root omega ↪ o;
+rule root (Cl $a) ↪ o;
+
+\end{lstlisting}
 
@@
 We may have notice that the lemmas in which the $comp$ symbol is used are only valid for a subset of formulas.
-The formulas need to have all of its quantifiers of sort $Graph$ and can only use the language $\in$, $\simeq$ and the classic logical connectives.
+The formulas need to have all of its quantifiers of sort $El~graph$ and can only use the language $\in$, $\simeq$ and the classic logical connectives.
 
@@
 
-Henceforth, we are able to define in \dedukti
-$$comp : \Graph \ra (\El \ nat \ \ra \ Graph) \  \ra \El \ formula \ra \Graph$$
-
 Then we are able to define an induction over formulas, using the language $\in$, $\simeq$ and the classic logical connectives.
@@
 The next step is to interpret an object of type $formula$ into $Prop$.
-We introduce the constant $interpretation$ which receives a valuation of type $\El \ nat \ \ra \ Graph$ and a formula of type $\El \ formula$ and return a $\Prop$.
+We introduce the constant $interpretation$ which receives a valuation of type $\El \ nat \ \ra \ \El~graph$ and a formula of type $\El \ formula$ and return a $\El~prop$.
 
 \begin{lstlisting}
-symbol interpretation : (El nat → Graph) → El formula → Prop;
+symbol interpretation : (El nat → El graph) → El formula → El prop;
 \end{lstlisting}
@@
 We need to have a tool to update a valuation when we assign a variable.
-To do so, we introduce the constant $update$ of type $(\El \ nat \ra Graph) \ra \El \ nat \ra Graph \ra (\El \ nat \ra Graph)$ which takes as arguments a valuation, a natural number and a graph and returns a new valuation: $(update \ \sigma \ x \ a)$ is a new valuation that substitues $x$ by $a$ and acts like $\sigma$ for the other natural numbers.
+To do so, we introduce the constant $update$ of type $(\El \ nat \ra \El~graph) \ra \El \ nat \ra \El~graph \ra (\El \ nat \ra \El~graph)$ which takes as arguments a valuation $\sigma$, a natural number $x$ and a graph $a$ and returns a new valuation $(update \ \sigma \ x \ a)$ that substitues $x$ by $a$ and acts like $\sigma$ for the other natural numbers.
 
@@
 
-We define the symbol $update1$ of type $(\El \ nat \ \ra \ Graph) \ \ra \ \El \ nat \ \ra \ Graph \ \ra \ \El \ nat \ \ra \ (\El \ nat \ \ra \ Graph)$.
-We keep in memory the argument of the valuation with the rewriting rule:
+We define the symbol $update1$ of type $(\El \ nat \ \ra \ \El~graph) \ \ra \ \El \ nat \ \ra \ \El~graph \ \ra \ \El \ nat \ \ra \ (\El \ nat \ \ra \ \El~graph)$.
+The new argument $z$ is used to keep in memory the argument $y$:
 
@@
 
-We have two natural numbers to compare: $x$, which is substitued by $a$ by the $update$, and $y$ that is the argument we apply to the valuation.
-The technique we use to compare $x$ and $y$ is the following: we keep in memory $y$ in the variable $z$, we decrement the two numbers until either one or both are equal to zero.
-If both are equal to zero, then $x$ and $y$ are equal and we return $a$.
-If only one equals zero, then they are different and we return $\sigma \ z$.
+We have two natural numbers to compare: $x$, which is substitued by $a$, and $y$ that is the argument we apply to the valuation. The technique we use to compare $x$ and $y$ is the following:
+
+\begin{itemize}
+\item We keep in memory $y$ in the variable $z$;
+\item We decrement $x$ and $y$ until either one or both are equal to zero;
+\item If both are equal to zero, then $x$ and $y$ are equal and we return $a$. If only one equals zero, then they are different and we return $\sigma \ z$.
+\end{itemize}
 
@@
 $$interpretation \ \sigma \ tF \lra \bltop$$
+
+\subsection{Comprehension and infinity}
+
+Henceforth, we are able to define in \dedukti
+$$comp : \El~graph \ra (\El \ nat \ \ra \ \El~graph) \  \ra \El \ formula \ra \ \El~graph$$
+
+and its rewriting rules
+
+\begin{lstlisting}
+rule eta (comp $a $σ $f) $x $x' ↪ 
+(`∃ y, `∃ y', (($x = i y) ∧ ($x' = i y') ∧ eta $a y y')) 
+∨ (`∃ y, ($x = i y) ∧ ($x' = o) ∧ (eta $a y (root $a))
+∧ (interpretation (update $σ zero (cr $a y)) $f));
+rule root (comp $a $σ $f) ↪ o;
+\end{lstlisting}
+
+When it comes to symbol related to the Infinity section of \cite{zermodulo}, we implement $empty_set$ of type $\El~graph$ and $Ind$ of type $\El~graph \ \ra \ \El~prop$.
+
+To define the empty set, we use $comp$ with the formula $fF$:
+\begin{lstlisting}
+rule empty_set ↪ comp omega (λ _, empty_set) fF;
+rule root empty_set ↪ o;
+\end{lstlisting}
+
+Then we implement
+
+\begin{lstlisting}
+rule Ind $c ↪ (empty_set ∈ $c) 
+∧ (`∀ a, (a ∈ $c) ⇒ ((join (pair a (pair a a))) ∈ $c));
+\end{lstlisting}
 
@@
 
-The second lemma is not required as it is a consequence of the rewriting rule of the polumorphic $=$ implemented in the stardard library of \dedukti: 
-\begin{lstlisting}
-constant symbol = {s} : El s → El s → Prop;
-notation = infix 4;
-rule π (@= $s $x $y) ↪ Π (P : El $s → Prop), π(P $x) → π(P $y);
-\end{lstlisting}
-
 \subsection{Useful mechanisms}
@@
 
-Lemma32
+Lemma32 + utilisation du lemma32
 
@@
 
-We want to deduce $weak \ extensionality$ from $strong \ extensionality$ (i.e. lemma 41).
+We notice in \cite{zermodulo53} the use of \textit{weak extensionality} to prove lemmas 44, 47 and 48.
+We want to deduce \textit{weak extensionality} from \textit{strong extensionality} (i.e.
+lemma 41).
 
-\begin{lstlisting}
-opaque symbol lemma41 : Π (x y : El nat), Π (f : El formula), Π (c d : Graph), π(`∀ σ : (El nat → Graph), ((interpretation (update (update σ x c) y d) f)
-∧ (`∀ a, `∀ a', `∀ b, ((a' ∈ a) ∧ (interpretation (update (update σ x a) y b) f)) ⇒ (`∃ b', ((b' ∈ b) ∧ (interpretation (update (update σ x a') y b') f))))
-∧ (`∀ b, `∀ b', `∀ a, ((b' ∈ b) ∧ (interpretation (update (update σ x a) y b) f)) ⇒ (`∃ a', ((a' ∈ a) ∧ (interpretation (update (update σ x a') y b') f))))) ⇒ (c simeq d))
-
-opaque symbol lemmaExt : Π (c d : Graph), π((`∀ x, (x ∈ c) ⇔ (x ∈ d)) ⇒ (c simeq d))
-\end{lstlisting}
-
-To do this, we need to prove the following intermediry lemma:
+To do this, we need to prove the following intermediary lemma:
 
@@
 \section{Conclusion}
+
+Schema pour représenter Z, Zmod et ZF ?
 

### 2021-12-02 theoriezf/Article/lambdapi.tex
Add files via upload

\usepackage{listings}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{xcolor}
\definecolor{green}{RGB}{0,130,0}
\definecolor{lightgrey}{RGB}{240,240,240}

\lstdefinelanguage{Lambdapi}
{
  inputencoding=utf8,
  extendedchars=true,
  numbers=none,
  numberstyle={},
  tabsize=2,
  basicstyle={\ttfamily\upshape\mdseries},
  backgroundcolor=\color{lightgrey},
  keywords={abort,admit,admitted,apply,as,assert,assertnot,associative,assume,begin,builtin,commutative,compute,constant,debug,end,fail,flag,focus,generalize,have,in,induction,inductive,infix,injective,left,let,notation,off,on,opaque,open,prefix,print,private,proofterm,protected,prover,prover_timeout,quantifier,refine,reflexivity,require,rewrite,right,rule,sequential,simplify,solve,symbol,symmetry,type,TYPE,unif_rule,verbose,why3,with},
  sensitive=true,
  keywordstyle=\color{blue},
  morecomment=[l]{//},
  commentstyle={\itshape\color{red}},
  string=[b]{"},
  stringstyle=\color{orange},
  showstringspaces=false,
  literate=
  {↪}{$\hookrightarrow$}1
  {→}{$\rightarrow$}1
  {Π}{$\Pi$}1
  {≔}{$\coloneqq$}1
  {𝔹}{$\mathbb{B}$}1
  {𝕃}{$\mathbb{L}$}1
  {ℕ}{$\mathbb{N}$}1
  {α}{$\alpha$}1
  {λ}{$\lambda$}1
  {σ}{$\sigma$}1
  {π}{$\pi$}1
  {τ}{$\tau$}1
  {ω}{$\omega$}1
  {∧}{$\wedge$}1
  {∨}{$\vee$}1
  {≤}{$\le$}1
  {≠}{$\neq$}1
  {∉}{$\notin$}1
  {×}{$\times$}1
  {ρ}{$\rho$}1
  {∈}{$\in$}1
  {∀}{$\forall$}1
  {∃}{$\exists$}1
  {⇒}{$\Rightarrow$}1
  {⇔}{$\Leftrightarrow$}1
  {⊤}{$\top$}1
  {⊥}{$\bot$}1
}
\lstset{language={Lambdapi}}

### 2021-12-16 theoriezf/Article/Article.tex
Add files via upload

@@
 
-%\usepackage{paralist, tabularx}
-\usepackage{oldlfont,amssymb,epsf,stmaryrd,epsfig,color}\usepackage{amsmath}
-%\usepackage{hyperref}
-%\usepackage{framed,multicol,changepage}
-%\usepackage{graphicx}
+\usepackage{amsmath}
 \usepackage[dvipsnames]{xcolor}
-%\usepackage{skull,ifoddpage,marginnote}
-%\usepackage{pifont}
-%\usepackage{graphicx}
-%\usepackage{ebproof}
-%\usepackage{mathtools}
-%\usepackage{bm}
-%\usepackage{stmaryrd}
-%\usepackage{cleveref}
 \usepackage{tikz}
-%\usepackage{import} % For inkscape generated files
 \usepackage{biblatex}
 \usepackage{url}
+\usepackage{fancyvrb}
+\usepackage{enumitem}
+\setitemize[1]{label=---}
 
-\title{Title}
+\title{\textbf{Dedukti: an implementation of set theory with pointed graphs} (titre provisoire)}
 \date{}
@@
 
+\begin{abstract}
+We show how to implement set theory in \dedukti ~in such a way that it respects normalization.
+To do so, we adapt in $\lambda \Pi$\textit{--calculus modulo theory} the guideline given in \textit{Deduction modulo} by A.
+Miquel and G.
+Dowek --using a theory of pointed graphs.
+To achieve this goal, we develop a language of formulas.
+\end{abstract}
+
 \section{Introduction}
 
-\dedukti \ is a logical framework based upon the language \lpcm, which was created to overcome the lack of interoperability between proof assistanrs.
-As it is a logical framework, we can implement in \dedukti \ theories.
-Especially, \textit{Set theory} is interesting beacause it will enable traductions from and towards \textsc{Isabelle/ZF}.
+\dedukti \ is a logical framework based upon the language \lpcm, which was created to overcome the lack of interoperability between proof assistants.
+As it is a logical framework, we can implement theories in \dedukti.
+Especially, set theory is interesting beacause it will enable traductions from and towards proof assistants using set theory, such as \textsc{Isabelle/ZF}.
 
-A previous work \cite{zermodulo} has demonstrate the implemtation of \textit{Set theory} in \textit{Deduction modulo} thaks to a structure of \textit{pointed graphs}.
-Until the writing of \cite{zermodulo}, \lpcm \ has been created as an improvement of \textit{Deduction modulo}.
+A previous work by A.
+Miquel and G.
+Dowek \cite{zermodulo} has demonstrated the implemtation of set theory in \textit{Deduction modulo} thaks to a structure of \textit{pointed graphs}.
+Since the writing of \cite{zermodulo}, \lpcm \ has been created as an improvement of \textit{Deduction modulo}.
 We adapt the implementation to the \lpcm \ and then implement it in \dedukti.
@@
 
-To implement $Set \ theory$ in \dedukti, the first idea is to translate the axioms of the theory into rewriting rules.
+To implement set theory in \dedukti, the first idea is to translate the axioms of the theory into rewriting rules.
 Nevertheless, as pointed out by M.
@@
 Miquel and G.
-Dowek \cite{zermodulo} introduced a theory of pointed graphs such that the implementation of $Set~theory$ respects normalization.
+Dowek \cite{zermodulo} introduced a structure of pointed graphs to set theory.
+This set theory with pointed graphs -- called IZmod -- respects normalization.
 
@@
 
-Sets are reprented by pointed graphs. A pointed graph is characterizd by a direct graph and one of its node identified as the root.
+In this theory, sets are reprented by pointed graphs. A pointed graph is a direct graph whom one of its node is identified as the root.
 
@@
 $a/x$ replace the root of $a$ by $x$.
-$x \eta_a y$ means there is a edge in $a$ from $y$ to $x$.
-$rel(x,y,r)$ means that $x$ and $y$ are related by the relation $r$.
+$x~\eta_a~y$ means there is a edge in $a$ from $y$ to $x$.
+$rel(x,y,r)$ means that the nodes $x$ and $y$ are related by the relation $r$.
+We have the following rewriting rules:
+$$ x \ \eta_{a/z} \ y \longrightarrow x \ \eta_a \ y $$
+$$ \text{root}(a/x) \longrightarrow x $$
+$$ (a/x)/y \longrightarrow a/y $$
 
-Because some different pointed graphs represents the same set, we introduce the bisimilarity:
+We saw in the examples that some different pointed graphs can represent the same set. To characterized this, A. Miquel and G. Dowek introduce bisimilarity between graphs (noted $\simeq$):
 
@@
 
-\subsection{The theory IZmod}
+\begin{figure}[h]
+\begin{tikzpicture}[>=stealth,yscale=1.8,xscale=1.4]
 
-IZmod
+\node(b) at (5,3)[circle,draw,text centered,fill=gray] {b};
+\node(c) at (5,2.5)[circle,draw,text centered] {c};
+\node(d) at (7,3)[circle,draw,text centered] {d};
+\node(e) at (6.5,2.5)[circle,draw,text centered] {e};
+\node(f) at (7.5,2.5)[circle,draw,text centered,fill=gray] {f};
+\node(g) at (7.5,2)[circle,draw,text centered] {g};
+\node(h) at (5.75,2.75)[text centered]{$\simeq$};
+\node(i) at (10,2.5)[text width=4cm,text centered] {The root is noted with the filled circle.};
+
+\draw[->] (b) -- (c); 
+\draw[->] (d) -- (e);
+\draw[->] (d) -- (f);
+\draw[->] (f) -- (g);
+
+\end{tikzpicture}
+\end{figure}
+
+The objective is to composed pointed graphs, for instance to join pointed graphs.
+The root of the new pointed graph is $o$.
+To guarantee that $o$ is not a node of one of the original pointed graphs, an injective function $i$ is introduced, such as its left inverse $i'$ and its image $I$.
+$$i'(i(x)) \lra x$$
+$$ I(x) \lra \top $$
+$$ I(o) \lra \bot $$
+
+The same method is applied to constructors $pair$, $powerset$, $etc$.
+
+\subsection{Set theories}
+
+Many definitions of the set theory intervines.
+The two classical set theories are Zermelo (Z) and Zermelo-Fraenkel (ZF).
+The set theory defined in \textit{Deduction modulo} by A.
+Miquel and G.
+Dowek \cite{zermodulo} is stronger than Zermelo set theory but weaker than Zermelo-Fraenkel set thery.
+Indeed, IZst contains Zermelo set theory, the Strong Extensionality axiom and the Transitive Closure axiom.
+\\
+
+\textbf{Strong Extensionality axiom.} $\fa~x_1,...,\fa~x_n,\fa~a,\fa~b, \\
+(R(a, b) \\
+\conj~\fa~x,\fa~x',\fa~y,(x' \in x \conj R(x, y) \imp \ex~y', (y' \in y \conj R(x', y'))) \\
+\conj \fa~y,\fa~y',\fa~x, (y' \in y \conj R(x, y) \imp \ex~x'(x' \in x \conj R(x', y'))) \\
+\imp a = b) $
+
+where $R(a,b)$ is a formula of free variables $x_1, ..., x_n$. \\
+
+\textbf{Transitive Closure axiom.} $\fa~a,\ex~e, (a \subseteq e \conj \fa~x,\fa~y, (x \in y \conj y \in e \imp x \in e))$ \\
+
+The Strong Extensionality axiom can be deduced from the Foundation axiom and the Transitive Closure axiom is can be derived from Replacement Scheme.
+
+The Strong Extensionality axiom implies the Extensionality axiom\footnote{See 5.3}.
+
+\begin{figure}[h]
+\begin{tikzpicture}[>=stealth,yscale=1,xscale=1]
+
+\draw[rounded corners=3pt] (0.5,0.5) rectangle (4,5);
+\draw[rounded corners=3pt] (0.25,0.25) rectangle (8,5.25);
+\draw[rounded corners=3pt] (0,0) rectangle (12,5.5);
+
+\node(a) at (2.25,3)[text width=3cm,text centered] {\textbf{Z} \\ Pairing \\ Union \\ Extensionality \\ Powerset \\ Comprehension \\ Infinity};
+\node(b) at (6,3)[text width=10cm,text centered] {\textbf{IZst} \\ Transitive Closure \\ Strong Extensionality};
+\node(c) at (10,3)[text width=3cm,text centered] {\textbf{ZF} \\ Replacement \\ Foundation};
+
+\end{tikzpicture}
+\caption{View of the different set theories}
+\end{figure}
 
@@
 
-So the signature is reduced to 26 symbols. The specific case of the comprehension symbol is treated later.
+The signature is thus reduced to 26 symbols. The specific case of the comprehension symbol is treated later.
 
@@
 
-Since the publication of \cite{zermodulo}, \lpcm has been developped as an evolution of \textit{Deduction modulo}.
+Since the publication of \cite{zermodulo}, \lpcm~ has been developped as an evolution of \textit{Deduction modulo}.
 
-More specifically, \lpcm allows to quantify on proposition.
+More specifically, \lpcm~ allows to quantify on proposition.
 
@@
 
+Similarly, the symbols $g_{x,y_1,...,y_n,P}$ and $g'_{x,x',y_1,...,y_n,P}$ are not necessary in \dedukti.
+These symbols were specific constructors that allows to introduce in $mem$ a proposition $P$ by convert it into a class.
+\\
+
+The rewriting rule $$y = z \ \lra \ \fa \ p \ (mem(y, p) \ \imp \ mem(z, p))$$ is not required as it is a $mem$ is useless in \lpcm.
+
 \subsection{Rewriting rules}
-
-The rewriting rule $$y = z \ \lra \ \fa \ p \ (mem(y, p) \ \imp \ mem(z, p))$$ is not required as it is a consequence of the rewriting rule of the polymorphic $=$ implemented in the stardard library of \dedukti:
-\begin{lstlisting}
-constant symbol = [s] : El s → El s → El prop;
-notation = infix 4;
-rule π (@= $s $x $y) ↪ Π (P : El $s → El prop), π(P $x) → π(P $y);
-\end{lstlisting}
-
-+ Expliquer pourquoi on n'utilise pas les symboles g et g' \\
 
@@
 
-\subsection{Comprehension and infinity}
-
-Henceforth, we are able to define in \dedukti
-$$comp : \El~graph \ra (\El \ nat \ \ra \ \El~graph) \  \ra \El \ formula \ra \ \El~graph$$
-
-and its rewriting rules
-
-\begin{lstlisting}
-rule eta (comp $a $σ $f) $x $x' ↪ 
-(`∃ y, `∃ y', (($x = i y) ∧ ($x' = i y') ∧ eta $a y y')) 
-∨ (`∃ y, ($x = i y) ∧ ($x' = o) ∧ (eta $a y (root $a))
-∧ (interpretation (update $σ zero (cr $a y)) $f));
-rule root (comp $a $σ $f) ↪ o;
-\end{lstlisting}
-
-When it comes to symbol related to the Infinity section of \cite{zermodulo}, we implement $empty_set$ of type $\El~graph$ and $Ind$ of type $\El~graph \ \ra \ \El~prop$.
-
-To define the empty set, we use $comp$ with the formula $fF$:
-\begin{lstlisting}
-rule empty_set ↪ comp omega (λ _, empty_set) fF;
-rule root empty_set ↪ o;
-\end{lstlisting}
-
-Then we implement
-
-\begin{lstlisting}
-rule Ind $c ↪ (empty_set ∈ $c) 
-∧ (`∀ a, (a ∈ $c) ⇒ ((join (pair a (pair a a))) ∈ $c));
-\end{lstlisting}
-
 \subsection{Results concerning valuation}
@@
 \begin{proof}
-The first term of the conjonction is proved by simple recurrence over natural numbers. The second term of the conjonction is proved by double recurrence.
+The first term of the conjonction is proved by simple recurrence over natural numbers. The second term of the conjonction is proved by double recurrence. $\square$
 \end{proof}
@@
 \begin{proof}
-The fifth theorem is proved by induction over formulas.
+The fifth theorem is proved by induction over formulas. $\square$
 \end{proof}
+
+\subsection{Comprehension and infinity}
+
+Henceforth, we are able to define in \dedukti
+$$comp : \El~graph \ra (\El \ nat \ \ra \ \El~graph) \  \ra \El \ formula \ra \ \El~graph$$
+
+and its rewriting rules
+
+\begin{lstlisting}
+rule eta (comp $a $σ $f) $x $x' ↪ 
+(`∃ y, `∃ y', (($x = i y) ∧ ($x' = i y') ∧ eta $a y y')) 
+∨ (`∃ y, ($x = i y) ∧ ($x' = o) ∧ (eta $a y (root $a))
+∧ (interpretation (update $σ zero (cr $a y)) $f));
+rule root (comp $a $σ $f) ↪ o;
+\end{lstlisting}
+
+When it comes to symbol related to the Infinity section of \cite{zermodulo}, we implement $empty\_set$ of type $\El~graph$ and $Ind$ of type $\El~graph \ \ra \ \El~prop$.
+
+To define the empty set, we use $comp$ with the formula $fF$:
+\begin{lstlisting}
+rule empty_set ↪ comp omega (λ _, empty_set) fF;
+rule root empty_set ↪ o;
+\end{lstlisting}
+
+Then we implement
+
+\begin{lstlisting}
+rule Ind $c ↪ (empty_set ∈ $c) 
+∧ (`∀ a, (a ∈ $c) ⇒ ((join (pair a (pair a a))) ∈ $c));
+\end{lstlisting}
 
@@
 
-\subsection{Useful mechanisms}
+The second lemma is already a consequence of the rewriting rule of the polymorphic $=$ implemented in the stardard library of \dedukti: 
+\begin{lstlisting}
+constant symbol = [s] : El s → El s → El prop;
+notation = infix 4;
+rule π (@= $s $x $y) ↪ Π (P : El $s → El prop), π(P $x) → π(P $y);
+\end{lstlisting}
 
-All the other lemmas except the ones where $comp$ is involved are proved using the blueprint in \cite{zermodulo53}.
+All the other lemmas of IZmod except the ones where $comp$ is involved are proved using the blueprint in \cite{zermodulo53}.
 The complete proofs can be found in \url{https://github.com/ttraversie/zf/tree/main/theoriezf}.
 
-\subsection{Some interesting proofs}
+\subsection{An example of proof}
 
-Lemma32 + utilisation du lemma32
+To show the way lemmas are proved in \dedukti ~we will take the example of lemma 30 and comment its proof. This lemma states that $$ a \in b \conj a \simeq c \imp c \in b $$
+
+\begin{proof}
+We first assume graphs $a$, $b$ and $c$ and $H$ the proof of $ a \in b \conj a \simeq c $. $a \in b$ rewrites to $\ex x~(x~\eta_b~root~b \conj a \simeq (b/x))$.
+
+We make appear $x$ and $Hx$ the proof $x~\eta_b~root~b \conj a \simeq (b/x)$.
+
+As the goal is to prove $c \in b$, that is to say $\ex y~(y~\eta_b~root~b \conj c \simeq (b/y))$, we need to find a suitable $y$.
+We take $x$ and now have two goals: $y~\eta_b~root~b$ and $c \simeq (b/y)$.
+
+The first one is proved by applying the left part of $Hx$.
+
+The second one is obtained by applying lemma 5 to $c$, $a$ and $b/x$.
+To apply lemma 5, we need to prove $c \simeq a \imp a \simeq b/x$.
+$c \simeq a$ is proved applying reflexivity to $a \simeq c$ (i.e.
+applying lemma 4 to $a$, $c$ and the right part of $H$).
+$a \simeq b/x$ derives from the right part of $Hx$.
+$\square$
+\end{proof}
+
+This proof is written in \dedukti thanks to the following code:
+
+\begin{lstlisting}
+opaque symbol lemma30 : π(`∀ a, `∀ b, `∀ c, 
+	((a ∈ b) ∧ (a simeq c)) ⇒ (c ∈ b))
+≔ begin
+assume a b c H;
+refine ExElim node _ (and_eg _ _ H) _ _;
+assume x Hx;
+refine ExIntro node x _ _;
+refine and_i _ _ _ _
+{refine and_eg _ _ Hx}
+{refine lemma5 c a (cr b x) 
+	(and_i _ _ (lemma4 a c (and_ed _ _ H)) (and_ed _ _ Hx))}
+end;
+\end{lstlisting}
+
+\subsection{Lemmas involving formulas}
+
+Now that the language of formulas have been designed along with the implementation of the $comp$ symbol, lemma 32 can be implemented.
+
+$$(P(z \leftarrow a) \conj a \simeq b) \imp P(z \leftarrow b)$$
+
+We implement it thanks to the $interpretation$ symbol. The valuation $update~\sigma~z~a)$ represents the assignement of variable $z \leftarrow a$.
+
+\begin{lstlisting}
+opaque symbol lemma32 : Π (z : El nat), Π (f : El formula), 
+π(`∀ a, `∀ b, (`∀ σ : (El nat → El graph),
+((interpretation (update σ z a) f) ∧ (a simeq b)) 
+⇒ (interpretation (update σ z b) f)))
+\end{lstlisting}
+
+The proof of this $opaque~symbol$ is done by induction over formulas: each case is proved easily, using the lemmas that have already been checked by \dedukti. \\
+
+Lemma 41 is implemented similarly:
+
+\begin{lstlisting}
+opaque symbol lemma41 : Π (x y : El nat), Π (f : El formula), 
+Π (c d : El graph), π(`∀ σ : (El nat → El graph), 
+((interpretation (update (update σ x c) y d) f)
+∧ (`∀ a, `∀ a', `∀ b, 
+	((a' ∈ a) 
+	∧ (interpretation (update (update σ x a) y b) f)) 
+	⇒ (`∃ b', ((b' ∈ b) 
+		∧ (interpretation (update (update σ x a') y b') f))))
+∧ (`∀ b, `∀ b', `∀ a, 
+	((b' ∈ b) 
+	∧ (interpretation (update (update σ x a) y b) f)) 
+	⇒ (`∃ a', ((a' ∈ a) 
+		∧ (interpretation (update (update σ x a') y b') f))))) 
+⇒ (c simeq d))
+\end{lstlisting}
 
@@
 
-We notice in \cite{zermodulo53} the use of \textit{weak extensionality} to prove lemmas 44, 47 and 48.
-We want to deduce \textit{weak extensionality} from \textit{strong extensionality} (i.e.
+We notice in \cite{zermodulo53} the use of weak extensionality (or simply extensionality) to prove lemmas 44, 47 and 48.
+We want to deduce weak extensionality from strong extensionality (i.e.
 lemma 41).
+\\
 
-To do this, we need to prove the following intermediary lemma:
+\textbf{Weak extensionality.} $\fa~c,\fa~d, (\fa~z, (z \in c \Leftrightarrow z \in d) \imp c \simeq d)$ \\
+
+We follow the blueprint given by G. Dowek and A. Miquel \cite{zermodulo} : we use the strong extensionality axiom where $R(x,y)$ is $(x \simeq c~\conj~y \simeq d)~\disj~x \simeq y$. \\
+
+\begin{proof}
+We want to prove that $\fa~c,\fa~d, (\fa~z, (z \in c \Leftrightarrow z \in d) \imp c \simeq d)$.
+We assume that $\fa~z, (z \in c \Leftrightarrow z \in d)$.
+We want to apply strong extensionality to deduce $c \simeq d$.
+To do so, we need to prove the three terms of the hypothesis of strong extensionality.
+
+$(c \simeq c~\conj~d \simeq d)~\disj~c \simeq d$ is a tautology.
+
+We want to prove that $\fa~a,\fa~a',\fa~b, (a' \in a~\conj~((a \simeq c~\conj~b \simeq d)~\disj~a \simeq b) \imp (\ex~b', (b' \in b~\conj~((a' \simeq c~\conj~b' \simeq d)~\disj~a' \simeq b'))$.
+We assume $(a' \in a~\conj~((a \simeq c~\conj~b \simeq d)~\disj~a \simeq b)$.
+If $(a \simeq c~\conj~b \simeq d)$, then we choose $b' \simeq a'$.
+We have $b' \in a$ because $a' \in a$.
+Yet, $a \simeq c$, $c \simeq d$ and $d \simeq b$.
+Then $b' \in b$.
+If $a \simeq b$, we choose $b' \simeq a'$.
+We have $b' \in a$ because $a' \in a$.
+Yet, $a \simeq b$.
+Thus $b' \in b$.
+
+We procede similarly for the third term. $\square$
+\end{proof}
+
+We impelement this theorem in \dedukti:
 
 \begin{lstlisting}
-opaque symbol lemmaHypExt : Π (c d : Graph), π((`∀ z, (z ∈ c) ⇔ (z ∈ d)) ⇒ ((((c simeq c) ∧ (d simeq d)) ∨ (c simeq d))
-∧ (`∀ a, `∀ a', `∀ b, ((a' ∈ a) ∧ (((a simeq c) ∧ (b simeq d)) ∨ (a simeq b))) ⇒ (`∃ b', ((b' ∈ b) ∧ (((a' simeq c) ∧ (b' simeq d)) ∨ (a' simeq b')))))
-∧ (`∀ b, `∀ b', `∀ a, ((b' ∈ b) ∧ (((a simeq c) ∧ (b simeq d)) ∨ (a simeq b))) ⇒ (`∃ a', ((a' ∈ a) ∧ (((a' simeq c) ∧ (b' simeq d)) ∨ (a' simeq b')))))))
+opaque symbol lemmaHypExt : Π (c d : Graph), 
+π((`∀ z, (z ∈ c) ⇔ (z ∈ d)) ⇒ 
+((((c simeq c) ∧ (d simeq d)) ∨ (c simeq d))
+∧ (`∀ a, `∀ a', `∀ b, 
+	((a' ∈ a) ∧ (((a simeq c) ∧ (b simeq d)) ∨ (a simeq b))) 
+	⇒ (`∃ b', ((b' ∈ b) 
+	∧ (((a' simeq c) ∧ (b' simeq d)) ∨ (a' simeq b')))))
+∧ (`∀ b, `∀ b', `∀ a, 
+	((b' ∈ b) ∧ (((a simeq c) ∧ (b simeq d)) ∨ (a simeq b)))
+	 ⇒ (`∃ a', ((a' ∈ a) 
+	 ∧ (((a' simeq c) ∧ (b' simeq d)) ∨ (a' simeq b')))))))
 \end{lstlisting}
 
-We thus can prove $weak \ extensionality$ thanks to:
+To prove $weak \ extensionality$, we assume graphs $c$ and $d$, and $H$ the hypothesis $\fa~z, (z \in c) \Leftrightarrow (z \in d)$. Then we apply lemma41 to:
+
+\begin{itemize}
+\item nats $zero$ and $one$
+\item the formula $(orF~(andF~(eqF~zero~two)~(eqF~one~three))~(eqF~zero~one))$
+\item graphs $c$ and $d$
+\item the valuation $(update~(update~(\lambda \_, empty\_set)~two~c)~three~d)$ 
+\item the proof of the left hand term $lemmaHypExt~c~d~H$.
+\end{itemize}
+
+Indeed, in the formula $$(orF~(andF~(eqF~zero~two)~(eqF~one~three))~(eqF~zero~one))$$ $zero$ and $one$ will be interpreted by $c$ and $d$ thanks to lemma41 and $two$ will be interpreted by $c$ and $three$ by $d$ thanks the valuation $$(update~(update~(\lambda \_, empty\_set)~two~c)~three~d)$$ The proposition obtained corresponds to the proposition in lemmaHypExt.
 
@@
 assume c d H;
-refine lemma41 zero one (orF (andF (eqF zero two) (eqF one three)) (eqF zero one)) c d (update (update (λ _, empty_set) two c) three d) (lemmaHypExt c d H);
+refine lemma41 zero one 
+		(orF (andF (eqF zero two) (eqF one three)) (eqF zero one)) 
+		c d 
+		(update (update (λ _, empty_set) two c) three d) 
+		(lemmaHypExt c d H);
 end;
@@
 
-Schema pour représenter Z, Zmod et ZF ?
+We have implemented in \dedukti ~a version of set theory -- IZst -- that respects normalization properties.
+IZst set theory correponds to Zermelo set theory, with Strong Extensionality axiom and Transitive Closure axiom.
+To do so, we have adapted the work by A.
+Miquel and G.
+Dowek \cite{zermodulo} from \textit{Deduction modulo} to \lpcm.
+We have \textit{formally} written all the proofs of the lemmas allowing us to implement set theory in \dedukti.
+
+To define and prove the lemmas corresponding to Comprehension axiom, we have developped a language of formulas, along with operators \textit{interpretation} and \textit{update}.
+In particular, the language of formulas allows us to prove that Weak Extensionality axiom derives from Strong Extensionality axiom.
 
@@
 \hline 17 & ? & 43 & ? \\
-\hline 18 & ? & 44 & ? \\
-\hline 19 & ? & 45 & ? \\
-\hline 20 & ? & 46 & ? \\
-\hline 21 & ? & 47 & ? \\
-\hline 22 & ? & 48 & ? \\
-\hline 23 & ? & 49 & ? \\
-\hline 24 & ? & 50 & ? \\
-\hline 25 & ? & 51 & ? \\
-\hline 26 & ? & 52 & ? \\
-\hline 27 & ? & 53 & ? \\
-\hline 28 & ? & Weak extensionality & ? \\
+\hline 18 & ? & Weak extensionality & ? \\
+\hline 19 & ? & 44 & ? \\
+\hline 20 & ? & 45 & ? \\
+\hline 21 & ? & 46 & ? \\
+\hline 22 & ? & 47 & ? \\
+\hline 23 & ? & 48 & ? \\
+\hline 24 & ? & 49 & ? \\
+\hline 25 & ? & 50 & ? \\
+\hline 26 & ? & 51 & ? \\
+\hline 27 & ? & 52 & ? \\
+\hline 28 & ? & 53 & ? \\
 \hline

### 2021-12-16 Article.tex
Add files via upload

--- 2021-12-16 theoriezf/Article/Article.tex
@@
 \usepackage{tikz}
-\usepackage{biblatex}
+\usepackage[backend=biber,style=alphabetic]{biblatex}
 \usepackage{url}
@@
 We show how to implement set theory in \dedukti ~in such a way that it respects normalization.
-To do so, we adapt in $\lambda \Pi$\textit{--calculus modulo theory} the guideline given in \textit{Deduction modulo} by A.
-Miquel and G.
-Dowek --using a theory of pointed graphs.
+To do so, we adapt in $\lambda \Pi$\textit{--calculus modulo theory} the guideline given in \textit{Deduction modulo} by G.
+Dowek and A.
+Miquel --using a theory of pointed graphs.
 To achieve this goal, we develop a language of formulas.
@@
 
-A previous work by A.
-Miquel and G.
-Dowek \cite{zermodulo} has demonstrated the implemtation of set theory in \textit{Deduction modulo} thaks to a structure of \textit{pointed graphs}.
+A previous work by G.
+Dowek and A.
+Miquel \cite{zermodulo} has demonstrated the implemtation of set theory in \textit{Deduction modulo} thaks to a structure of \textit{pointed graphs}.
 Since the writing of \cite{zermodulo}, \lpcm \ has been created as an improvement of \textit{Deduction modulo}.
@@
 Crabbé \cite{crabbé}, such implementation does not guarantee normalization.
-That is why A.
-Miquel and G.
-Dowek \cite{zermodulo} introduced a structure of pointed graphs to set theory.
+That is why G.
+Dowek and A.
+Miquel \cite{zermodulo} introduced a structure of pointed graphs to set theory.
 This set theory with pointed graphs -- called IZmod -- respects normalization.
@@
 
-We saw in the examples that some different pointed graphs can represent the same set. To characterized this, A. Miquel and G. Dowek introduce bisimilarity between graphs (noted $\simeq$):
+We saw in the examples that some different pointed graphs can represent the same set. To characterized this, G. Dowek and A. Miquel introduce bisimilarity between graphs (noted $\simeq$):
 
@@
 The two classical set theories are Zermelo (Z) and Zermelo-Fraenkel (ZF).
-The set theory defined in \textit{Deduction modulo} by A.
-Miquel and G.
-Dowek \cite{zermodulo} is stronger than Zermelo set theory but weaker than Zermelo-Fraenkel set thery.
+The set theory defined in \textit{Deduction modulo} by G.
+Dowek and A.
+Miquel \cite{zermodulo} is stronger than Zermelo set theory but weaker than Zermelo-Fraenkel set thery.
 Indeed, IZst contains Zermelo set theory, the Strong Extensionality axiom and the Transitive Closure axiom.
@@
 
-The Strong Extensionality axiom implies the Extensionality axiom\footnote{See 5.3}.
+The Strong Extensionality axiom implies the Extensionality axiom\footnote{See \ref{extensionality}}.
 
@@
 
+\label{extensionality}
+
 We notice in \cite{zermodulo53} the use of weak extensionality (or simply extensionality) to prove lemmas 44, 47 and 48.
@@
 IZst set theory correponds to Zermelo set theory, with Strong Extensionality axiom and Transitive Closure axiom.
-To do so, we have adapted the work by A.
-Miquel and G.
-Dowek \cite{zermodulo} from \textit{Deduction modulo} to \lpcm.
+To do so, we have adapted the work by G.
+Dowek and A.
+Miquel \cite{zermodulo} from \textit{Deduction modulo} to \lpcm.
 We have \textit{formally} written all the proofs of the lemmas allowing us to implement set theory in \dedukti.

### 2021-12-19 theoriezf/Article/Article.tex
Add files via upload

@@
 \begin{abstract}
-We show how to implement set theory in \dedukti ~in such a way that it respects normalization.
-To do so, we adapt in $\lambda \Pi$\textit{--calculus modulo theory} the guideline given in \textit{Deduction modulo} by A.
-Miquel and G.
-Dowek --using a theory of pointed graphs.
+We show how to implement a version of set theory in \dedukti ~in such a way that it respects normalization.
+To do so, we adapt in $\lambda \Pi$\textit{--calculus modulo theory} the guideline given in \textit{Deduction modulo} by G.
+Dowek and A.
+Miquel --using a theory of pointed graphs-- and we \textit{formally} write the proofs in \dedukti.
 To achieve this goal, we develop a language of formulas.
@@
 To implement set theory in \dedukti, the first idea is to translate the axioms of the theory into rewriting rules.
-Nevertheless, as pointed out by M.
+However, as pointed out by M.
 Crabbé \cite{crabbé}, such implementation does not guarantee normalization.
-That is why A.
-Miquel and G.
-Dowek \cite{zermodulo} introduced a structure of pointed graphs to set theory.
+That is why G.
+Dowek and A.
+Miquel \cite{zermodulo} introduced a structure of pointed graphs to encode set theory.
 This set theory with pointed graphs -- called IZmod -- respects normalization.
 
-\subsection{Pointed graphs}
+\subsection{Sets as pointed graphs}
 
-In this theory, sets are reprented by pointed graphs. A pointed graph is a direct graph whom one of its node is identified as the root.
+\label{informal}
+
+In the IZmod theory \cite[see][Section 3.1]{zermodulo}, sets are reprented by pointed graphs. A pointed graph is a direct graph whose one of its node is identified as the root.
 
 \begin{figure}[h]
-\begin{tikzpicture}[>=stealth,yscale=1.8,xscale=1.4]
+\centering
+\begin{tikzpicture}[>=stealth,yscale=1.8,xscale=1.35]
 
-\node(a) at (3,3)[circle,draw,text centered] {a};
-\node(b) at (5,3)[circle,draw,text centered] {b};
-\node(c) at (5,2.5)[circle,draw,text centered] {c};
-\node(d) at (7,3)[circle,draw,text centered] {d};
-\node(e) at (6.5,2.5)[circle,draw,text centered] {e};
-\node(f) at (7.5,2.5)[circle,draw,text centered] {f};
-\node(g) at (7.5,2)[circle,draw,text centered] {g};
+\node(h1) at (3,3.5)[text centered]{$\emptyset$};
+\node(a) at (3,3)[circle,draw,text centered,fill=gray] {a};
+\node(h2) at (4,3.5)[text centered]{$\{\emptyset\}$};
+\node(b) at (4,3)[circle,draw,text centered,fill=gray] {b};
+\node(c) at (4,2.5)[circle,draw,text centered] {c};
+\node(h3) at (5,3.5)[text centered]{$\emptyset$};
+\node(b1) at (5,3)[circle,draw,text centered] {b};
+\node(c1) at (5,2.5)[circle,draw,text centered,fill=gray] {c};
+\node(h3) at (6.5,3.5)[text centered]{$\{\emptyset,\{\emptyset\}\}$};
+\node(d) at (6.5,3)[circle,draw,text centered,fill=gray] {d};
+\node(e) at (6,2.5)[circle,draw,text centered] {e};
+\node(f) at (7,2.5)[circle,draw,text centered] {f};
+\node(g) at (7,2)[circle,draw,text centered] {g};
+\node(h4) at (8.5,3.5)[text centered]{$\emptyset$};
+\node(d1) at (8.5,3)[circle,draw,text centered] {d};
+\node(e1) at (8,2.5)[circle,draw,text centered,fill=gray] {e};
+\node(f1) at (9,2.5)[circle,draw,text centered] {f};
+\node(g1) at (9,2)[circle,draw,text centered] {g};
+\node(h5) at (10.5,3.5)[text centered]{$\{\emptyset\}$};
+\node(d2) at (10.5,3)[circle,draw,text centered] {d};
+\node(e2) at (10,2.5)[circle,draw,text centered] {e};
+\node(f2) at (11,2.5)[circle,draw,text centered,fill=gray] {f};
+\node(g2) at (11,2)[circle,draw,text centered] {g};
+\node(h6) at (12.5,3.5)[text centered]{$\emptyset$};
+\node(d3) at (12.5,3)[circle,draw,text centered] {d};
+\node(e3) at (12,2.5)[circle,draw,text centered] {e};
+\node(f3) at (13,2.5)[circle,draw,text centered] {f};
+\node(g3) at (13,2)[circle,draw,text centered,fill=gray] {g};
 
 \draw[->] (b) -- (c); 
+\draw[->] (b1) -- (c1); 
 \draw[->] (d) -- (e);
@@
 \draw[->] (f) -- (g);
+\draw[->] (d1) -- (e1);
+\draw[->] (d1) -- (f1);
+\draw[->] (f1) -- (g1);
+\draw[->] (d2) -- (e2);
+\draw[->] (d2) -- (f2);
+\draw[->] (f2) -- (g2);
+\draw[->] (d3) -- (e3);
+\draw[->] (d3) -- (f3);
+\draw[->] (f3) -- (g3);
 
@@
 
-The first graph with root $a$ represents the set $\emptyset$.
-The second graph with root $b$ represents the set $\{\emptyset\}$, while with root c it represents $\emptyset$.
-The third graph with root $d$ represents $\{\emptyset,\{\emptyset\}\}$, \textit{etc}.
+We have represented several pointed graphs where the root is indicated by the filled circle.
+We see that the location of the root is important as it changes the set represented by the pointed graph.
+In the third graph, the node $b$ has no effect because the root is $c$.
+\\
 
@@
 
-We saw in the examples that some different pointed graphs can represent the same set. To characterized this, A. Miquel and G. Dowek introduce bisimilarity between graphs (noted $\simeq$):
+We see in the examples that some different pointed graphs can represent the same set.
+To characterized this, G.
+Dowek and A.
+Miquel introduce bisimilarity between graphs, noted $\simeq$, and with rewriting rule
 
-$a \simeq b \lra \ex r~(rel(root(a),root(b),r) \\
-\conj \ \fa x \fa x'~\fa y~(x'~\eta_a~x~\conj~rel(x, y, r)~\imp~\ex~y'~(y'~\eta_b~y~\conj~rel(x', y', r))) \\
-\conj \ \fa y \fa y'~\fa x~(y'~\eta_b~y~\conj~rel(x, y, r)~\imp~\ex~x'~(x'~\eta_a~x~\conj~rel(x', y', r))))$
+\begin{equation*}
+\begin{split}
+a \simeq b \lra~ &\ex~r~(rel(root(a),root(b),r) \\
+&\conj \ \fa x \fa x'~\fa y~(x'~\eta_a~x~\conj~rel(x, y, r)~\imp~\ex~y'~(y'~\eta_b~y~\conj~rel(x', y', r))) \\
+&\conj \ \fa y \fa y'~\fa x~(y'~\eta_b~y~\conj~rel(x, y, r)~\imp~\ex~x'~(x'~\eta_a~x~\conj~rel(x', y', r))))
+\end{split}
+\end{equation*}
 
 \begin{figure}[h]
+\centering
 \begin{tikzpicture}[>=stealth,yscale=1.8,xscale=1.4]
@@
 \node(h) at (5.75,2.75)[text centered]{$\simeq$};
-\node(i) at (10,2.5)[text width=4cm,text centered] {The root is noted with the filled circle.};
 
@@
 \end{figure}
+
+In set theory, there only exists one sort: the sets.
+In the IZmod theory, there exists two sorts: nodes and pointed graphs.
+Therefore, there are in IZmod two constructions related to the inclusion of sets.
+We can take the example of $\emptyset \in \{\emptyset\}$.
+We can represents this inclusion by two different ways.
+
+\begin{figure}[h]
+\centering
+\begin{tikzpicture}[>=stealth,yscale=1.8,xscale=1.4]
+
+\node(a) at (3,2.75)[circle,draw,text centered,fill=gray] {h};
+\node(k) at (3.5,2.75)[text centered]{$\in$};
+\node(b) at (4,3)[circle,draw,text centered,fill=gray] {i};
+\node(c) at (4,2.5)[circle,draw,text centered] {j};
+\node(d) at (6,3)[circle,draw,text centered] {k};
+\node(e) at (6,2.5)[circle,draw,text centered] {$\ell$};
+
+\draw[->] (b) -- (c); 
+\draw[->] (d) -- (e);
+
+\end{tikzpicture}
+\end{figure}
+
+Into a pointed graphs, inclusion is represented by the constructor $\eta$.
+Indeed, there exists an edge from $k$ to $\ell$, which means that the pointed graph with root $i$ is included in the pointed graph with root $h$.
+
+Between two pointed graphs, inclusion is represented by a new constructor noted $\in$, with rewriting rule $a \in b \lra \ex~x~(x~\eta_b~root(b)~\conj~a \simeq (b/x))$. \\
 
@@
 The two classical set theories are Zermelo (Z) and Zermelo-Fraenkel (ZF).
-The set theory defined in \textit{Deduction modulo} by A.
-Miquel and G.
-Dowek \cite{zermodulo} is stronger than Zermelo set theory but weaker than Zermelo-Fraenkel set thery.
-Indeed, IZst contains Zermelo set theory, the Strong Extensionality axiom and the Transitive Closure axiom.
+The set theory defined in \textit{Deduction modulo} by G.
+Dowek and A.
+Miquel \cite[see][Section 2]{zermodulo} --called IZst-- is stronger than Zermelo set theory but weaker than Zermelo-Fraenkel set thery.
+Indeed, the IZst theory contains Zermelo set theory, the Strong Extensionality axiom and the Transitive Closure axiom.
 \\
 
-\textbf{Strong Extensionality axiom.} $\fa~x_1,...,\fa~x_n,\fa~a,\fa~b, \\
-(R(a, b) \\
-\conj~\fa~x,\fa~x',\fa~y,(x' \in x \conj R(x, y) \imp \ex~y', (y' \in y \conj R(x', y'))) \\
-\conj \fa~y,\fa~y',\fa~x, (y' \in y \conj R(x, y) \imp \ex~x'(x' \in x \conj R(x', y'))) \\
-\imp a = b) $
+\textbf{Strong Extensionality axiom.} 
+\begin{equation*}
+\begin{split}
+\fa~x_1,...,\fa~x_n,\fa~a,\fa~b, &(R(a, b) \\
+&\conj~\fa~x,\fa~x',\fa~y,(x' \in x \conj R(x, y) \imp \ex~y', (y' \in y \conj R(x', y'))) \\
+&\conj~\fa~y,\fa~y',\fa~x, (y' \in y \conj R(x, y) \imp \ex~x'(x' \in x \conj R(x', y'))) \\
+&\imp a = b) 
+\end{split}
+\end{equation*}
 
@@
 
+The hypothesis of the Strong Extensionality axiom imitates the structure of the rewriting rule of $\eta$.
+This can be explained with the following reasoning: in the Extensionality axiom, the hypothesis states that two sets have the same elements, so the hypothesis of the strong extensionality axiom copy the structure of one of the equivalent to the inclusion in IZmod --that is to say the constructor $eta$.
+\\
+
 \textbf{Transitive Closure axiom.} $\fa~a,\ex~e, (a \subseteq e \conj \fa~x,\fa~y, (x \in y \conj y \in e \imp x \in e))$ \\
+
+The Transitive Closure axiom conveys the idea that every set is included in a transitive set. \\
 
@@
 
-The Strong Extensionality axiom implies the Extensionality axiom\footnote{See 5.3}.
+The Strong Extensionality axiom implies the Extensionality axiom\footnote{See \ref{extensionality}}.
+
+The IZmod theory is an extension of IZst set theory with an encoding of pointed graphs.
 
 \begin{figure}[h]
+\centering
 \begin{tikzpicture}[>=stealth,yscale=1,xscale=1]
@@
 
-The langage of the theory IZmod uses four sorts. The first two are for
+The langage of the theory IZmod uses four sorts \cite[see][Section 3.2]{zermodulo}. The first two are for
 the pointed graphs and for the nodes of the pointed graphs.  In
@@
 The symbols $graph$ and $node$ are specific to the
-expression of $IZmod$ in \dedukti. In contrast the symbols $\Set$,
+expression of IZmod in \dedukti. In contrast the symbols $\Set$,
 $\El$, $prop$, and $\arr$ are part of the standard library of 
@@
 
-The signature of $IZmod$ contains 31 symbols. As we have replaced the
+The signature of IZmod contains 31 symbols \cite[see][Table 2]{zermodulo}. As we have replaced the
 sorts for classes and relations with the \dedukti \ types
@@
 
-Since the publication of \cite{zermodulo}, \lpcm~ has been developped as an evolution of \textit{Deduction modulo}.
+Since the publication of \cite{zermodulo}, \lpcm ~has been developped as an evolution of \textit{Deduction modulo}.
+More specifically, \lpcm ~allows to quantify on proposition.
+In that respect, the schemes of axioms have been replaced by axioms, as henceforth we do not need to write one axiom for each proposition.
+\\
 
-More specifically, \lpcm~ allows to quantify on proposition.
+In \dedukti, the symbol $mem$ and $rel$ \cite[see][Table 2]{zermodulo} are useless. Indeed, we can apply a node $x$ to class $P$ with $P~x$ and two nodes $x$ and $y$ to a relation $r$ with $r~x~y$.
 
-In \dedukti, the symbol $mem$ and $rel$ are useless. Indeed, we can apply a node $x$ to class $P$ with $P~x$ and two nodes $x$ and $y$ to a relation $r$ with $r~x~y$.
-
-Similarly, the symbols $g_{x,y_1,...,y_n,P}$ and $g'_{x,x',y_1,...,y_n,P}$ are not necessary in \dedukti.
-These symbols were specific constructors that allows to introduce in $mem$ a proposition $P$ by convert it into a class.
+The symbols $g_{x,y_1,...,y_n,P}$ and $g'_{x,x',y_1,...,y_n,P}$ had been created as specific constructors that allow to introduce in $mem$ a proposition $P$ by convert it into a class.
+These symbols are now not necessary in \dedukti since $mem$ is deleted.
 \\
@@
 
-All the other rewriting rules, except the ones involving $comp$, are easy to implement in \dedukti.
+All the other rewriting rules \cite[see][Table 3]{zermodulo}, except the ones involving $comp$, are easy to implement in \dedukti.
+
+\begin{center}
+\underline{\textbf{General}}
+\end{center}
 
@@
 rule (cr (cr $a $x) $y) ↪ cr $a $y;
+\end{lstlisting}
 
+\begin{center}
+\underline{\textbf{Relocations}}
+\end{center}
+
+\begin{lstlisting}
 rule i' (i $x) ↪ $x;
@@
 rule $x < (Succ $y) ↪ ($x < $y) ∨ ($x = $y);
+\end{lstlisting}
 
+\begin{center}
+\underline{\textbf{Equality and membership}}
+\end{center}
+
+\begin{lstlisting}
 rule $a simeq $b ↪ `∃ r : El (node arrow (node arrow prop)), 
@@
 rule $a ∈ $b ↪ `∃ x, ((eta $b x (root $b)) ∧ ($a simeq cr $b x));
+\end{lstlisting}
 
+\begin{center}
+\underline{\textbf{Constructions}}
+\end{center}
+
+\begin{lstlisting}
 rule eta (join $a) $x $x' ↪ 
@@
 
+The pointed graph $omega$ represents the set of Von Neumann ordinals \cite[see][Table 2]{zermodulo53}.
+Indeed, the informal presentation of the representation of sets by pointed graphs given in \ref{informal} corresponds to Von Neumann ordinals.
+The rewriting rule of $x~\eta_{omega}~x'$ states that in $omega$ the edges represents the order relation $<$ and that there is an edge from every natural number to the root of $omega$.
+
 \section{The language of formulas}
 
-We may have notice that the lemmas in which the $comp$ symbol is used are only valid for a subset of formulas.
+We may have notice that the lemmas in which the $comp$ symbol is used are only valid for a subset of formulas \cite[see][Table 5]{zermodulo}.
 The formulas need to have all of its quantifiers of sort $El~graph$ and can only use the language $\in$, $\simeq$ and the classic logical connectives.
@@
 
-\subsection{Comprehension and infinity}
+\subsection{Comprehension, empty set and inductive set}
 
@@
 
-When it comes to symbol related to the Infinity section of \cite{zermodulo}, we implement $empty\_set$ of type $\El~graph$ and $Ind$ of type $\El~graph \ \ra \ \El~prop$.
+When it comes to the symbol related to the Infinity section of \cite[see][Section 2.1]{zermodulo}, we implement $empty\_set$ of type $\El~graph$ and $Ind$ of type $\El~graph \ \ra \ \El~prop$.
 
@@
 
+There are 53 lemmas to proved \textit{formally} \cite[see][Tables a and 5]{zermodulo}.
+
 The first lemma $x=x$ does not need to be implemented since it is already part of the standard library of \dedukti \ under the name $refl$ (which is polymorphic).
@@
 
-All the other lemmas of IZmod except the ones where $comp$ is involved are proved using the blueprint in \cite{zermodulo53}.
+All the other lemmas of IZmod except the ones where $comp$ is involved are proved using the blueprint in \cite[see][Proposition 1]{zermodulo53}.
 The complete proofs can be found in \url{https://github.com/ttraversie/zf/tree/main/theoriezf}.
@@
 
-We notice in \cite{zermodulo53} the use of weak extensionality (or simply extensionality) to prove lemmas 44, 47 and 48.
+\label{extensionality}
+
+We notice in \cite{zermodulo53} the use of weak extensionality (simply called extensionality) to prove lemmas 44, 47 and 48.
 We want to deduce weak extensionality from strong extensionality (i.e.
@@
 
+\subsection{The axioms of IZst theory}
+
+We have now encoded in \dedukti ~all the axioms of IZst set theory: the strong extensionality axiom corresponds to lemma 41, the axiom of the union is implemented by lemma 42, the pairing axiom corresponds to lemma 43, the axiom of the power set is encoded by lemma 44, the comprehension scheme is implemented by lemma 45, the axiom of infinity corresponds to lemma 51 and the transitive closure axiom is encoded by lemmas 52 and 53.
+
 \section{Conclusion}
@@
 We have implemented in \dedukti ~a version of set theory -- IZst -- that respects normalization properties.
-IZst set theory correponds to Zermelo set theory, with Strong Extensionality axiom and Transitive Closure axiom.
-To do so, we have adapted the work by A.
-Miquel and G.
-Dowek \cite{zermodulo} from \textit{Deduction modulo} to \lpcm.
+IZst set theory correponds to Zermelo set theory, with the Strong Extensionality axiom and the Transitive Closure axiom.
+To do so, we have adapted the work by G.
+Dowek and A.
+Miquel \cite{zermodulo} from \textit{Deduction modulo} to \lpcm.
 We have \textit{formally} written all the proofs of the lemmas allowing us to implement set theory in \dedukti.
 
-To define and prove the lemmas corresponding to Comprehension axiom, we have developped a language of formulas, along with operators \textit{interpretation} and \textit{update}.
-In particular, the language of formulas allows us to prove that Weak Extensionality axiom derives from Strong Extensionality axiom.
+To define and prove the lemmas corresponding to the Comprehension axiom, we have developped a language of formulas, along with operators \textit{interpretation} and \textit{update}.
+In particular, the language of formulas allows us to prove that the Extensionality axiom derives from the Strong Extensionality axiom.
 

### 2022-01-13 theoriezf/Article/Article.tex
Add files via upload

@@
 We show how to implement a version of set theory in \dedukti ~in such a way that it respects normalization.
-To do so, we adapt in $\lambda \Pi$\textit{--calculus modulo theory} the guideline given in \textit{Deduction modulo} by G.
+To do so, we adapt in $\lambda \Pi$\textit{--calculus modulo theory} the guideline given in \textit{Deduction modulo theory} by G.
 Dowek and A.
@@
 
-\dedukti \ is a logical framework based upon the language \lpcm, which was created to overcome the lack of interoperability between proof assistants.
-As it is a logical framework, we can implement theories in \dedukti.
-Especially, set theory is interesting beacause it will enable traductions from and towards proof assistants using set theory, such as \textsc{Isabelle/ZF}.
+Several prover are based on set theory, such as \textsc{Mizar}, \textsc{Atelier B} and \textsc{Isabelle/ZF}.
+The goal of this paper is to present an implementation of set theory in the logical framework \dedukti ~\cite{expressing} expressed in the \lpcm.
+We use here one of \dedukti's implementation, called \textsc{Lambdapi}.
+Our long term project is to use this implementation to enable interoperability between such provers based on set theory.
 
-A previous work by G.
-Dowek and A.
-Miquel \cite{zermodulo} has demonstrated the implemtation of set theory in \textit{Deduction modulo} thaks to a structure of \textit{pointed graphs}.
-Since the writing of \cite{zermodulo}, \lpcm \ has been created as an improvement of \textit{Deduction modulo}.
-We adapt the implementation to the \lpcm \ and then implement it in \dedukti.
+The first way to implement the set theory in \dedukti ~is to state each axiom in the framework.
+But such a method does not yield a cut-elimination theorem for the set theory.
+In particular, it does not enable to extract witnesses from a constructive existence proof.
+An alternative is to orient these axioms as rewriting rules.
+For instance, $$x \in \mathcal{P}(y) \Leftrightarrow x \subseteq y$$ is replaced with $$x \in \mathcal{P}(y) \lra x \subseteq y$$
+
+However, as pointed out by M.
+Crabbé \cite{crabbé}, such a formulation of set theory does not have a cut-elimination property.
+An other solution is to define sets as \textit{pointed graphs}.
+With such a formulation, we can prove a cut-elimination theorem: for every proof in natural deduction that contains cuts there exists a cut-free proof.
+Such a theory expressed in \textit{Dedution modulo theory} is presented in \cite{zermodulo} with pencil and paper proofs.
+The goal of this paper is to present an implementation of this theory.
+
+In 2007, the \lpcm ~had not been defined and this is why the original theory was defined in \textit{Dedution modulo theory}.
+The first issue was to adapt the expression in \textit{Dedution modulo theory} to the expression in \lpcm.
+This led to simplify the original formulation as classes of nodes of pointed graphs and relations on nodes can be naturaly expressed in \dedukti ~while that required a specific sort and specific comprehension axiom in \textit{Dedution modulo theory}.
+This enabled to reduce the size of the signature from 31 symbols to 26.
+
+On the other hand, in the original formulation of the theory some lemmas were only valid for a specific subset of propositions.
+These lemmas were proved by induction over the structure of the propositions of this subset.
+Implementing these lemmas in \dedukti ~required to define a new inductive sort for the formulas of this subset and an intepretation of these formulas into the general type of  propositions, itself defined with rewriting rules.
+This development is the first, to our knowledge, to use such a reflexion principle defined with rewriting rules in \dedukti.
+The generality of this method still remains to be investigated.
 
@@
 
-To implement set theory in \dedukti, the first idea is to translate the axioms of the theory into rewriting rules.
-However, as pointed out by M.
-Crabbé \cite{crabbé}, such implementation does not guarantee normalization.
-That is why G.
+The set theory with pointed graphs developped by G.
 Dowek and A.
-Miquel \cite{zermodulo} introduced a structure of pointed graphs to encode set theory.
-This set theory with pointed graphs -- called IZmod -- respects normalization.
+Miquel is called IZmod (Intuitionistic Zermelo set theory in \textit{Deduction modulo}).
+We give here an informal presentation of the ideas developped in \cite{zermodulo}.
 
@@
 
-In the IZmod theory \cite[see][Section 3.1]{zermodulo}, sets are reprented by pointed graphs. A pointed graph is a direct graph whose one of its node is identified as the root.
+In the IZmod theory, sets are reprented by pointed graphs. A pointed graph is a direct graph whose one of its node is identified as the root.
 
@@
 $x~\eta_a~y$ means there is a edge in $a$ from $y$ to $x$.
-$rel(x,y,r)$ means that the nodes $x$ and $y$ are related by the relation $r$.
+The notation $rel(x,y,r)$ means that the nodes $x$ and $y$ are related by the relation $r$.
 We have the following rewriting rules:
 $$ x \ \eta_{a/z} \ y \longrightarrow x \ \eta_a \ y $$
-$$ \text{root}(a/x) \longrightarrow x $$
+$$ root(a/x) \longrightarrow x $$
 $$ (a/x)/y \longrightarrow a/y $$
 
-We see in the examples that some different pointed graphs can represent the same set.
-To characterized this, G.
-Dowek and A.
-Miquel introduce bisimilarity between graphs, noted $\simeq$, and with rewriting rule
+We see in the examples that some different pointed graphs can represent the same set. To characterized this, a notion of  bisimilarity, noted $\simeq$, is introduced between graphs with rewriting rule
 
@@
 \begin{split}
-a \simeq b \lra~ &\ex~r~(rel(root(a),root(b),r) \\
-&\conj \ \fa x \fa x'~\fa y~(x'~\eta_a~x~\conj~rel(x, y, r)~\imp~\ex~y'~(y'~\eta_b~y~\conj~rel(x', y', r))) \\
-&\conj \ \fa y \fa y'~\fa x~(y'~\eta_b~y~\conj~rel(x, y, r)~\imp~\ex~x'~(x'~\eta_a~x~\conj~rel(x', y', r))))
+a \simeq b \lra~ &\ex~r~root(a)~root(b) \\
+&\conj \ \fa x \fa x'~\fa y~(x'~\eta_a~x~\conj~r~x~y~\imp~\ex~y'~(y'~\eta_b~y~\conj~r~x'~y')) \\
+&\conj \ \fa y \fa y'~\fa x~(y'~\eta_b~y~\conj~r~x~y~\imp~\ex~x'~(x'~\eta_a~x~\conj~r~x'~y')))
 \end{split}
@@
 We can take the example of $\emptyset \in \{\emptyset\}$.
-We can represents this inclusion by two different ways.
+We can represents this inclusion by two different ways, one between pointed graphs and the other between nodes within a pointed graph.
 
@@
 
-\node(a) at (3,2.75)[circle,draw,text centered,fill=gray] {h};
-\node(k) at (3.5,2.75)[text centered]{$\in$};
-\node(b) at (4,3)[circle,draw,text centered,fill=gray] {i};
-\node(c) at (4,2.5)[circle,draw,text centered] {j};
 \node(d) at (6,3)[circle,draw,text centered] {k};
-\node(e) at (6,2.5)[circle,draw,text centered] {$\ell$};
+\node(e) at (6,2.5)[circle,draw,text centered,fill=gray] {$\ell$};
+\node(k) at (6.5,2.75)[text centered]{$\in$};
+\node(f) at (7,3)[circle,draw,text centered,fill=gray] {k};
+\node(g) at (7,2.5)[circle,draw,text centered] {$\ell$};
 
-\draw[->] (b) -- (c); 
 \draw[->] (d) -- (e);
+\draw[->] (f) -- (g);
 
@@
 
-Into a pointed graphs, inclusion is represented by the constructor $\eta$.
-Indeed, there exists an edge from $k$ to $\ell$, which means that the pointed graph with root $i$ is included in the pointed graph with root $h$.
-
-Between two pointed graphs, inclusion is represented by a new constructor noted $\in$, with rewriting rule $a \in b \lra \ex~x~(x~\eta_b~root(b)~\conj~a \simeq (b/x))$. \\
+When in a graph we have an edge from $k$ to $\ell$, the pointed graph with the root $\ell$ represents a set which is an element of the set represented by the pointed graph with root $k$.
+But any pointed graph bisimilar with pointed graph with root $\ell$ also represents a set that is an element of the set represented with the pointed graph with root $k$.
+This leads to the definition of membership noted $\in$, with rewriting rule $a \in b \lra \ex~x~(x~\eta_b~root(b)~\conj~a \simeq (b/x))$.
+\\
 
@@
 The root of the new pointed graph is $o$.
-To guarantee that $o$ is not a node of one of the original pointed graphs, an injective function $i$ is introduced, such as its left inverse $i'$ and its image $I$.
+To guarantee that $o$ is not a node of one of the original pointed graphs, an injective function $i$ is introduced, with $i'$ its left inverse and $I$ the predicate of its image.
+
 $$i'(i(x)) \lra x$$
-$$ I(x) \lra \top $$
-$$ I(o) \lra \bot $$
+$$I(x) \lra \top$$
+$$I(o) \lra \bot$$
 
@@
 
-Many definitions of the set theory intervines.
-The two classical set theories are Zermelo (Z) and Zermelo-Fraenkel (ZF).
-The set theory defined in \textit{Deduction modulo} by G.
-Dowek and A.
-Miquel \cite[see][Section 2]{zermodulo} --called IZst-- is stronger than Zermelo set theory but weaker than Zermelo-Fraenkel set thery.
-Indeed, the IZst theory contains Zermelo set theory, the Strong Extensionality axiom and the Transitive Closure axiom.
+These pointed graphs permit to prove the axioms of a theory IZst that lies between Zermelo (Z) and Zermelo-Fraenkel (ZF).
+This theory does not contain the Replacement scheme but it contains two additional axioms: the Strong Extensionality axiom and the Transitive Closure axiom.
 \\
@@
 The hypothesis of the Strong Extensionality axiom imitates the structure of the rewriting rule of $\eta$.
-This can be explained with the following reasoning: in the Extensionality axiom, the hypothesis states that two sets have the same elements, so the hypothesis of the strong extensionality axiom copy the structure of one of the equivalent to the inclusion in IZmod --that is to say the constructor $eta$.
+This can be explained with the following reasoning: in the Extensionality axiom, the hypothesis states that two sets have the same elements, so the hypothesis of the strong extensionality axiom copy the structure of one of the equivalent to the inclusion in IZmod --that is to say the constructor $\eta$.
 \\
@@
 
-The Strong Extensionality axiom implies the Extensionality axiom\footnote{See \ref{extensionality}}.
+The Strong Extensionality axiom implies the Extensionality axiom\footnote{See Section \ref{extensionality}}.
 
@@
 
-\node(a) at (2.25,3)[text width=3cm,text centered] {\textbf{Z} \\ Pairing \\ Union \\ Extensionality \\ Powerset \\ Comprehension \\ Infinity};
-\node(b) at (6,3)[text width=10cm,text centered] {\textbf{IZst} \\ Transitive Closure \\ Strong Extensionality};
-\node(c) at (10,3)[text width=3cm,text centered] {\textbf{ZF} \\ Replacement \\ Foundation};
+\node(a) at (2.25,3)[text width=3cm,text centered] {\underline{\textbf{Z}} \\ Pairing \\ Union \\ Extensionality \\ Powerset \\ Comprehension \\ Infinity};
+\node(b) at (6,3)[text width=10cm,text centered] {\underline{\textbf{IZst}} \\ Transitive Closure \\ Strong Extensionality};
+\node(c) at (10,3)[text width=3cm,text centered] {\underline{\textbf{ZF}} \\ Replacement \\ Foundation};
 
@@
 
-\subsection{From \textit{Deduction modulo} to \lpcm}
+\subsection{From \textit{Deduction modulo theory} to \lpcm}
 
-Since the publication of \cite{zermodulo}, \lpcm ~has been developped as an evolution of \textit{Deduction modulo}.
-More specifically, \lpcm ~allows to quantify on proposition.
-In that respect, the schemes of axioms have been replaced by axioms, as henceforth we do not need to write one axiom for each proposition.
-\\
+Unlike the \textit{Dedution modulo theory}, the \lpcm ~allows quantification over propositions. \\
 
-In \dedukti, the symbol $mem$ and $rel$ \cite[see][Table 2]{zermodulo} are useless. Indeed, we can apply a node $x$ to class $P$ with $P~x$ and two nodes $x$ and $y$ to a relation $r$ with $r~x~y$.
+Thus the symbols $mem$ and $rel$ \cite[see][Table 2]{zermodulo} are not needed anymore.
+Indeed, a class of nodes can simply be expressed as a term $P$ of type $El~n \ra El~prop$, and the fact that a node $x$ is an element of this class can be expressed as $P~x$ and not as $mem(x,P)$.
+In a similar way, the fact that two nodes $x$ and $y$ are related by a relation $r$ is expressed with $r~x~y$.
 
 The symbols $g_{x,y_1,...,y_n,P}$ and $g'_{x,x',y_1,...,y_n,P}$ had been created as specific constructors that allow to introduce in $mem$ a proposition $P$ by convert it into a class.
-These symbols are now not necessary in \dedukti since $mem$ is deleted.
-\\
-
-The rewriting rule $$y = z \ \lra \ \fa \ p \ (mem(y, p) \ \imp \ mem(z, p))$$ is not required as it is a $mem$ is useless in \lpcm.
+These symbols are now not necessary in \dedukti ~since $mem$ is deleted.
 
@@
 
-All the other rewriting rules \cite[see][Table 3]{zermodulo}, except the ones involving $comp$, are easy to implement in \dedukti.
+All the other rewriting rules \cite[see][Table 3]{zermodulo}, except the ones involving the symbol associated with the construction of sets by comprehension (noted $comp$), are easy to implement in \dedukti.
 
@@
 
-The pointed graph $omega$ represents the set of Von Neumann ordinals \cite[see][Table 2]{zermodulo53}.
+The pointed graph \texttt{omega} represents the set of Von Neumann ordinals \cite[see][Table 2]{zermodulo53}.
 Indeed, the informal presentation of the representation of sets by pointed graphs given in \ref{informal} corresponds to Von Neumann ordinals.
-The rewriting rule of $x~\eta_{omega}~x'$ states that in $omega$ the edges represents the order relation $<$ and that there is an edge from every natural number to the root of $omega$.
+The rewriting rule of \texttt{eta omega x x'} states that in \texttt{omega} the edges represents the order relation $<$ and that there is an edge from every natural number to the root of \texttt{omega}.
 
@@
 
-We may have notice that the lemmas in which the $comp$ symbol is used are only valid for a subset of formulas \cite[see][Table 5]{zermodulo}.
+A key point in the original formulation of the theory is that the $comp$ symbol enabling to construct a set in comprehension is used are only valid for a subset of formulas \cite[see][Table 5]{zermodulo}.
 The formulas need to have all of its quantifiers of sort $El~graph$ and can only use the language $\in$, $\simeq$ and the classic logical connectives.
+Taking this restriction into account was the main challenge of this implementation.
 
@@
 
-There are 53 lemmas to proved \textit{formally} \cite[see][Tables a and 5]{zermodulo}.
+A first result on the original formulation of the theory was to prove that the axioms of IZst are theorems in IZmod.
+This required 53 lemmas that were \textit{informally} proved in \cite[see][Tables 4 and 5]{zermodulo} and that we proved \textit{formally} in this paper.
 
-The first lemma $x=x$ does not need to be implemented since it is already part of the standard library of \dedukti \ under the name $refl$ (which is polymorphic).
+Some of these proofs just follow the informal ones. Some others required the use of the type of formula and its embedding into propositions.
+
+The first lemma $x=x$ does not need to be implemented since it is already part of the standard library of \dedukti \ under the name \texttt{refl} (which is polymorphic).
 
@@
 
-This proof is written in \dedukti thanks to the following code:
+This proof is written in \dedukti ~thanks to the following code:
 
@@
 
-We implement it thanks to the $interpretation$ symbol. The valuation $update~\sigma~z~a)$ represents the assignement of variable $z \leftarrow a$.
+We implement it thanks to the $interpretation$ symbol. The valuation $update~\sigma~z~a$ represents the assignement of variable $z \leftarrow a$.
 
@@
 
-The proof of this $opaque~symbol$ is done by induction over formulas: each case is proved easily, using the lemmas that have already been checked by \dedukti. \\
+The proof of this \texttt{opaque symbol} is done by induction over formulas: each case is proved easily, using the lemmas that have already been checked by \dedukti. \\
 
@@
 
-We follow the blueprint given by G. Dowek and A. Miquel \cite{zermodulo} : we use the strong extensionality axiom where $R(x,y)$ is $(x \simeq c~\conj~y \simeq d)~\disj~x \simeq y$. \\
+We follow the blueprint given by G.
+Dowek and A.
+Miquel \cite[see][Proposition 1]{zermodulo} : we use the strong extensionality axiom where $R(x,y)$ is $(x \simeq c~\conj~y \simeq d)~\disj~x \simeq y$.
+\\
 
@@
 Dowek and A.
-Miquel \cite{zermodulo} from \textit{Deduction modulo} to \lpcm.
+Miquel \cite{zermodulo} from \textit{Deduction modulo theory} to \lpcm ~and have encoded set with a theory of pointed graphs -- IZmod.
 We have \textit{formally} written all the proofs of the lemmas allowing us to implement set theory in \dedukti.
@@
 In particular, the language of formulas allows us to prove that the Extensionality axiom derives from the Strong Extensionality axiom.
+
+The implementation of IZmod theory represents a significant corpus of formal proofs in \textsc{Lambdapi}.
+
+\newpage
+\section*{Annex}
 
